---
title: "Abundance ~ biogeographic model results - depth excluded"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    #fig_height: 
    fig_width: 140/25.4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```

# Load packages
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/extras.R")
```

# Load data
```{r}
load("data/Rdata/predsum.Rdata")
load(file = "data/Rdata/abun_biogeo_model_nodep_final.Rda")
load("data/Rdata/rmetadata.Rdata")
```

# Model investigation 

## Conditional effects plot again to aid interpretation
```{r}
abun_biogeo_model_nodep_final |> conditional_effects() |> plot(ask = FALSE, points = FALSE) # Without raw data
#abun_biogeo_model_nodep_final |> conditional_effects() |> plot(ask = FALSE, points = TRUE) # With raw data overplotted
```

## Summary table
```{r}
abun_biogeo_model_nodep_final |> 
  as_draws_df() |>    # Extract draws
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  rhat, ess_bulk, ess_tail,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1))|> 
  as_tibble() |> 
  mutate(median = round(median, 3),
         lower = round(lower, 3),
         upper = round(upper, 3))
```

## Summary
No interaction or any effect of dCT or dGBR


# If there were interactions, run the following, otherwise skip to the individual, marginal effects plots

Find max and min values for dCT and dGBR
```{r}
rmetadata |> 
  summarise(min_dCT = min(dCT),
            max_dCT = max(dCT),
            min_dGBR = min(dGBR),
            max_dGBR = max(dGBR))
```


```{r}
rmetadata |> 
  summarise(median = median(dCT), Q1 = quantile(dCT, probs = 0.25), Q4 = quantile(dCT, probs = 0.95),
            CI_05 = quantile(dCT, probs = 0.05), CI_95 = quantile(dCT, probs = 0.99))
```





# Make dataframe with data points:
```{r}
newdata <- data.frame(dCT = c(25, 1156),
                      dGBR = c(48, 355))
```


```{r}
abun_biogeo_model_nodep_final %>% 
  emmeans(~ dCT|dGBR, at = newdata, type = "response") %>% 
  as_tibble()

```



```{r}
interact_data1 <- predsum |> 
  with(list(dCT = seq(100, 1000, len = 100),
            dGBR = c(50, 350)))
```



# If there were an interaction you could run this:

```{r fig.width = 160/25.4, fig.height=55/25.4}
interaction_plot1 <- 
abun_biogeo_model_nodep_final |> 
  emmeans(~ dCT|dGBR, type = 'link', 
          epred = TRUE, 
          at = interact_data1) |> 
  regrid()  |> 
  gather_emmeans_draws()  |> 
  as_tibble() |>
  mutate(dGBRf = as_factor(case_when(dGBR ==50 ~ "50km dGBR",
                          dGBR == 350 ~ "350km dGBR"))) |> 
  ggplot(aes(x = dCT, y = .value, fill = dGBRf)) +
  stat_lineribbon(colour = "black", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.5) +
  stat_lineribbon(colour = "black", size = 0.5, point_interval = "median_hdci", .width = 0.8, alpha = 1) +
  scale_fill_brewer(type = "qual", palette = 'Accent')+
  theme_bjc() +
  scale_x_continuous("Distance from Coral Triangle (km)") +
  scale_y_continuous("Est. no preds. 150m2", expand = c(0.05,0)) +
  theme(axis.title = element_text(colour = "black", size = 10, family = "Arial"),
        axis.text = element_text(colour = "black", size = 9, family = "Arial"),
        strip.text = element_text(colour = "black", size = 10, family = "Arial"),
        legend.position = "none") +
  facet_wrap(~dGBRf)


interaction_plot1
```


```{r}
interact_data2 <- predsum |> 
  with(list(dGBR = seq(50, 350, len = 100),
            dCT = c(250, 750)))
```



```{r fig.width = 160/25.4, fig.height=55/25.4}
interaction_plot2 <- 
  abun_biogeo_model_nodep_final |> 
  emmeans(~ dCT|dGBR, type = 'link', 
          epred = TRUE, 
          at = interact_data2) |> 
  regrid()  |> 
  gather_emmeans_draws()  |> 
  as_tibble() |> 
  mutate(dCTf = case_when(dCT ==250 ~ "250km dCT",
                          dCT == 750 ~ "750km dCT")) |> 
  ggplot(aes(x = dGBR, y = .value, fill = dCTf)) +
  stat_lineribbon(colour = "black", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.5) +
  stat_lineribbon(colour = "black", size = 0.5, point_interval = "median_hdci", .width = 0.8, alpha = 1) +
  theme_bjc() +
  scale_fill_brewer(type = "qual", palette = 'Set2')+
  scale_x_continuous("Distance from GBR (km)") +
  scale_y_continuous("Est. no preds. 150m2", expand = c(0.05,0)) +
  theme(axis.title = element_text(colour = "black", size = 10, family = "Arial"),
        axis.text = element_text(colour = "black", size = 9, family = "Arial"),
        strip.text = element_text(colour = "black", size = 10, family = "Arial"),
        legend.position = "none") +
  facet_wrap(~dCTf)

interaction_plot2
```


# Combine
```{r fig.width = 160/25.4, fig.height=100/25.4}
dCT_dGBR_abun_fig <- 
  cowplot::plot_grid(interaction_plot1, interaction_plot2, 
                     nrow = 2, labels = c("a", "b"), rel_heights = c(1, 1), label_size = 10, align = "h") +
  theme(plot.background = element_rect(color = "black"))
dCT_dGBR_abun_fig
```


# Save
```{r}
#ggsave(filename = "output/dCT_dGBR_abun_fig.pdf", plot = dCT_dGBR_abun_fig, width = 160, height = 100, device = cairo_pdf, units = "mm", dpi = 300)
```




### Run these if no interaction ###


# Marginal effects plots

# dCT

# Make new data grid
```{r}
dCT_grid <- predsum |> 
  with(list(dCT = seq_range(dCT, n = 100)))
```


# Trend ribbon plot
```{r fig.height=60/26.5, fig.width=96/25.4}
dCT_abun_trend <- 
abun_biogeo_model_nodep_final |> 
  emmeans(~ dCT|dGBR, type = 'link', 
          epred = TRUE, 
          at = dCT_grid) |> 
  regrid()  |> 
  gather_emmeans_draws()  |> 
  as_tibble() |> 
  ggplot(aes(x = dCT, y = .value)) +
  stat_lineribbon(colour = "black", fill = "#C65894", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.5) +
  stat_lineribbon(colour = "black", fill = "#C65894", size = 0.5, point_interval = "median_hdci", .width = 0.8, alpha = 1) +
  theme_bjc() +
  scale_x_continuous("Distance from Coral Triangle (km)") +
  scale_y_continuous("No. preds (.150m2)", expand = c(0.05, 0)) +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
 #       strip.text = element_text(colour = "black", size = 9),
        #strip.text = element_blank(),
        plot.title = element_text(size = 10, hjust = 0.5),
        legend.position = "none") 

dCT_abun_trend
```


# Effect size density plot
```{r fig.width=56/25.4, fig.height=60/25.4}
dCT_abun_effect <- 
  abun_biogeo_model_nodep_final |> 
  gather_draws(`^b_.*`, regex=TRUE)  |> 
  filter(.variable == "b_scaledCT")   |> 
  ungroup() %>% 
  ggplot(aes(x = .value)) +
  stat_halfeye(.width = c(0.8, 0.95), fill = "#C65894", point_interval = "median_hdi") +
  theme_bjc() +
  scale_y_discrete(position = "right", expand = c(0.05, 0)) +
  annotate(geom = "text", label = "P=0.53", x = -0.4, y = 1, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5), limits = c(-0.5, 0.5)) +
  geom_vline(xintercept = 0, lty = 11) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(colour = "black", size = 10),
        axis.text.y.right = element_text(colour = "black", angle = 270, hjust = 1.2, size = 10),
        axis.text.x = element_text(colour = "black", size = 9),
        strip.text = element_blank()) +
  guides(fill = "none") +
  labs(x = "Scaled effect distribution")
dCT_abun_effect
```

```{r fig.width = 150/25.4, fig.height=55/25.4}
dCT_abun_fig <- 
  cowplot::plot_grid(dCT_abun_trend, dCT_abun_effect, 
                     nrow = 1, labels = c("a", "b"), rel_widths = c(1.2, 0.6), label_size = 10, align = "h") +
  theme(plot.background = element_rect(color = "black"))
dCT_abun_fig
```
# Save
```{r}
ggsave(filename = "output/dCT_abun_fig.pdf", plot = dCT_abun_fig, height = 60, width = 152, device = cairo_pdf, units = "mm", dpi = 300)
```


### dGBR

# Make new data grid
```{r}
dGBR_grid <- predsum |> 
  with(list(dGBR = seq_range(dGBR, n = 100)))
```


## Plot
```{r fig.height=60/26.5, fig.width=96/25.4}
dGBR_abun_trend <- 
  abun_biogeo_model_nodep_final |> 
  emmeans(~ dGBR, type = 'link',
          epred = TRUE,
           at = dGBR_grid)  |> 
  regrid()  |> 
  gather_emmeans_draws()  |> 
  as_tibble() %>% 
  ggplot(aes(x = dGBR, y = .value)) +
  stat_lineribbon(colour = "black", fill = "#FFCD00", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.5) +
  stat_lineribbon(colour = "black", fill = "#FFCD00", size = 0.5, point_interval = "median_hdci", .width = 0.85, alpha = 1) +
  #scale_colour_viridis_d(option = "plasma", end = 0.8) +
  #scale_fill_viridis_d(option = "plasma", end = 0.8) +
  theme_bjc() +
  scale_x_continuous("Distance from GBR (km)", limits = c(48, 355)) +
  scale_y_continuous("No. preds (.150m2)", expand = c(0.05, 0)) +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
 #       strip.text = element_text(colour = "black", size = 9),
        #strip.text = element_blank(),
        plot.title = element_text(size = 10, hjust = 0.5),
        legend.position = "none") 

dGBR_abun_trend 
```


# Effect size density plot
```{r fig.width=56/25.4, fig.height=60/25.4}
dGBR_abun_effect <- 
  abun_biogeo_model_nodep_final |> 
  gather_draws(`^b_.*`, regex=TRUE)  |> 
  filter(.variable == "b_scaledGBR")   |> 
  ungroup() %>% 
  ggplot(aes(x = .value)) +
  stat_halfeye(.width = c(0.85, 0.95), fill = "#FFCD00", point_interval = "median_hdi") +
  theme_bjc() +
  scale_y_discrete(position = "right", expand = c(0.05, 0)) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5), limits = c(-0.5, 0.5)) +
  annotate(geom = "text", label = "P=0.64", x = -0.4, y = 1, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  geom_vline(xintercept = 0, lty = 11) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(colour = "black", size = 10),
        axis.text.y.right = element_text(colour = "black", angle = 270, hjust = 1.2, size = 10),
        axis.text.x = element_text(colour = "black", size = 9),
        strip.text = element_blank()) +
  guides(fill = "none") +
  labs(x = "Scaled effect distribution")
dGBR_abun_effect
```

```{r fig.width = 150/25.4, fig.height=55/25.4}
dGBR_abun_fig <- 
  cowplot::plot_grid(dGBR_abun_trend, dGBR_abun_effect, 
                     nrow = 1, labels = c("a", "b"), rel_widths = c(1.2, 0.6), label_size = 10, align = "h") +
  theme(plot.background = element_rect(color = "black"))
dGBR_abun_fig
```



# Save
```{r}
ggsave(filename = "output/dGBR_abun_fig.pdf", plot = dGBR_abun_fig, height = 60, width = 152, device = cairo_pdf, units = "mm", dpi = 300)
```


```{r}
# END #
```







