---
title: "Abundance ~ Depth"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Introduction

Here we are subsetting just Bougainville, Osprey and Lihou reefs and modelling abundance over depth



# Load predsum.sub data
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
#sourceRmd("code/abundance/diw_abund.Rmd")
load("data/Rdata/predsum.sub.Rdata")
```







# Subset
```{r}
predsum.sub <- predsum |> 
  filter(Reef_1 %in% c("Lihou", "Osprey", "Bougainville")) %>% 
  droplevels()
```


# Check
```{r}
predsum.sub
```



# EDA

```{r}
predsum.sub |> 
  ggplot(aes(x = Depth, y = No_Preds)) +
  geom_boxplot() 
```


Lots of zeros. Will need ZI term in model


## Set model formula
```{r}
depth_form <- bf(No_Preds ~ Depth + (1|Site), zi ~ 1, family = zero_inflated_negbinomial(link = "log"))
```

## Check what priors we need to estimate
```{r}
get_prior(depth_form, data = predsum.sub)
```




## Use data properties to help set weakly informative priors
```{r}
predsum.sub |> 
  group_by(Depth) |> 
  summarise(median = median(No_Preds), MAD = mad(No_Preds), mean = mean(No_Preds), sd = sd(No_Preds), logmedian = log(median(No_Preds)), 
            logMAD = log(mad(No_Preds)), logmean = log(mean(No_Preds)), logsd = log(sd(No_Preds)))
```


## Set priors
```{r}
depth_priors <-  
  prior(normal(1, 2), class = 'Intercept') +
  prior(normal(0, 0.5), class = 'b') +
  prior(gamma(2, 2), class = 'shape') +
  prior(logistic(0, 1), class = 'Intercept', dpar = "zi")
```



## Fit model
```{r}
depth_mod1 <- 
  brm(depth_form,
  data = predsum.sub,
  prior = depth_priors,
  sample_prior = 'only',              
  chains = 3,
  cores = 3,
  iter = 5000, 
  warmup = 1000, 
  thin = 4,
  seed = 123,
  silent = 0,
  refresh = 0,
  backend = "cmdstanr")

```


## Check prior width
```{r fig.width=140/25.4}
depth_mod1 |> conditional_effects() |> plot(ask = FALSE, points = TRUE)
```

Looks like they should be wide enough. Will fit the data and do pp checks


# Fit model with data
```{r}
depth_mod2 <- depth_mod1 |> update(sample_prior = "yes", cores = 3, refresh = 0)
```


# Check influence of priors
```{r}
depth_mod2 |> SUYR_prior_and_posterior()
```


# Model validation

## DHARMa residuals
```{r}
resids <- make_brms_dharma_res(depth_mod2, integerResponse = FALSE)
testUniformity(resids)
plotResiduals(resids, form = factor(rep(1, nrow(predsum.sub))))
plotResiduals(resids, quantreg = TRUE)
testDispersion(resids) # Dispersion plot (dispersion = variance/mean) - we assume mean and variance are the same, so we need to check
```


## Chain checks

```{r}
# Trace plot
mcmc_plot(depth_mod2, type='trace')
stan_trace(depth_mod2$fit)
# Autocorrelation
mcmc_plot(depth_mod2, type='acf_bar')
stan_ac(depth_mod2$fit)
# Rhat - how well the chains converged
mcmc_plot(depth_mod2, type='rhat_hist')
# Percentage effectives - needs to be greater than 0.5
mcmc_plot(depth_mod2, type='neff_hist')

# Others
mcmc_plot(depth_mod2, type='dens_overlay')
mcmc_plot(depth_mod2, type='nuts_acceptance')
mcmc_plot(depth_mod2, type='nuts_energy')
```



## PP checks
```{r}
# Link scale
pp_check(depth_mod2, ndraws = 250)

# Transformed
pred <- posterior_predict(depth_mod2)
bayesplot::ppc_dens_overlay(y = log1p(predsum.sub$No_Preds), 
                            yrep = log1p(pred[1:100,]))
```
y = the observed data
yrep = the predictive(posterior) distribution
These follow approximate trends, so we'll go with it for now.


## From this model we should be able to work out the expected proportion of observations (transects) that would contain no predatory fishes


# Check R2
```{r}
depth_mod2 |> performance::r2_posterior() |> as.data.frame() |>
  median_hdci()
```


# Final model
```{r}
depth_mod_final <- depth_mod2
#save(depth_mod_final, file = "data/Rdata/depth_mod_final.Rdata")
```





# Model investigation 
## Conditional effects plot again to aid interpretation
```{r}
depth_mod2 |> conditional_effects() |> plot(ask = FALSE, points = FALSE) # Without raw data
#depth_mod2 |> conditional_effects() |> plot(ask = FALSE, points = TRUE) # With raw data overplotted
```


## Summary table
```{r}
depth_mod2 |> 
  as_draws_df() |>    # Extract draws
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  rhat, ess_bulk, ess_tail,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1))|> 
  as_tibble() |> 
  mutate(median = round(median, 3),
         lower = round(lower, 3),
         upper = round(upper, 3))
```


# Simple contrasts:
```{r}
depth_mod2 |> 
  emmeans(~Depth, type = "response")

depth_mod2 |> 
  emmeans(~Depth, regrid = TRUE)

depth_mod2 |> 
  emmeans(~Depth, type = "link") |> 
  regrid()
```


# Exceedence probablities
```{r}
depth_mod2 |> 
  emmeans(~Depth, type = "link") |> 
  regrid() |> 
  pairs()
```





```{r}
depth_mod2 |> 
  emmeans(~Depth, type = "link")  |> 
  gather_emmeans_draws() |> 
  ggplot(aes(x = Depth, y = .value)) +
 # geom_point(data = predsum %>% filter(No_Preds != 0), aes(x = Reef_1_Area, y = No_Preds), fill = "#66cc99", colour = "#336633", pch = 21, alpha = 0.3, position = position_jitter()) +
  geom_pointrange(colour = "black", fill = "#ffa494", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.6) +
  stat_lineribbon(colour = "black", fill = "#fa2600", size = 0.5, point_interval = "median_hdci", .width = 0.85, alpha = 0.9) +
  #scale_colour_viridis_d(option = "plasma", end = 0.8) +
#  scale_fill_manual(values = websafe_pinkpal) +
#  scale_fill_viridis_d(option = "plasma", end = 0.8) +
  #scale_y_continuous(limits = c(0,15), breaks = c(2,4,6,8,10,12)) +
  labs(x = "Reef area (kha)", 
       y = "No. preds (n.150m2)") +
  theme_minimal() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        strip.text = element_text(colour = "black", size = 9),
        plot.title = element_text(size = 10, hjust = 0.5),
        legend.position = "none")


```












