---
title: "Abundance ~ biogeographic model results"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    #fig_height: 
    fig_width: 140/25.4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```

# Load packages
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/extras.R")
```

# Load data
```{r}
load("data/Rdata/predsum.Rdata")
load(file = "data/Rdata/abun_biogeo_model_nodep_final.Rda")
```

# Model investigation 

## Conditional effects plot again to aid interpretation
```{r}
abun_biogeo_model_nodep_final |> conditional_effects() |> plot(ask = FALSE, points = FALSE) # Without raw data
#abun_biogeo_model_nodep_final |> conditional_effects() |> plot(ask = FALSE, points = TRUE) # With raw data overplotted
```

## Summary table
```{r}
abun_biogeo_model_nodep_final |> 
  as_draws_df() |>    # Extract draws
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  rhat, ess_bulk, ess_tail,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1))|> 
  as_tibble() |> 
  mutate(median = round(median, 3),
         lower = round(lower, 3),
         upper = round(upper, 3))
```

## Summary
Quite a complicated set of interactions here.

In the shallows at average dCT and average dGBR there were just over 3 predators per transect (95% CI [2.072 - 4.622])
The effect of depth varied depending on dCT and dGBR


## Probably need a couple of figs here 



# Make new data grid
```{r}
dCTgrid <- predsum |> 
  with(list(dCT = c(150, 500, 850),
            dGBR = median(dGBR)))
```



# dCT and Depth
```{r fig.height=80/26.5, fig.width=150/25.4}
dCT_depth_abun_plot <- 
abun_biogeo_model_nodep_final |> 
  emmeans(~ Depth|dCT, type = 'link', at = biogeogrid) |> 
  regrid()  |> 
  as_tibble() |> 
  mutate(dCTkm = paste(dCT, "km", sep = "")) %>% 
  ggplot(aes(x = Depth, y = prob)) +
  geom_line(aes(group = 1)) +
  geom_pointrange(aes(ymin = lower.HPD, ymax = upper.HPD, fill = Depth), pch = 21, size = 1, position = position_dodge(width = 0.2)) +
  scale_shape_manual(values = c(21,22,24)) +
  scale_fill_manual(values = websafe_bluepal) +
  #scale_colour_manual(values = c("#006ad1", "#1e90ff", "#6bb6ff")) +
  theme_bjc() +
  facet_wrap(~dCTkm) +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
 #       strip.text = element_text(colour = "black", size = 9),
        #strip.text = element_blank(),
        plot.title = element_text(size = 10, hjust = 0.5),
        legend.position = "none") 
dCT_depth_abun_plot
```

# Save
```{r}
ggsave(filename = "output/dCT_depth_abun_fig.pdf", plot = dCT_depth_abun_plot, height = 80, width = 150, device = cairo_pdf, units = "mm", dpi = 300)
```


# Plot depth emmeans and  contrasts at mean dCT and dGBR


# emmeans
```{r fig.height=80/26.5, fig.width=70/25.4}
#depth_em <- 
  abun_biogeo_model_nodep_final |> 
  emmeans(~ Depth, type = "response") |> 
  regrid()  |> 
  as_tibble() |> 
  ggplot(aes(x = Depth, y = prob)) +
  geom_pointrange(aes(ymin = lower.HPD, ymax = upper.HPD, fill = Depth), pch = 21, size = 1) +
  scale_fill_manual(values = websafe_bluepal) +
  scale_y_continuous("Abundance (n preds.150m2)") +
  theme_bjc() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        plot.title = element_text(size = 10, hjust = 0.5),
        legend.position = "none") 
depth_em
```


## Contrasts
```{r fig.height=80/26.5, fig.width=80/25.4}
depth_contr <- 
  abun_biogeo_model_nodep_final |> 
  emmeans(~ Depth, type = "response") |> 
 # regrid()  |> 
  pairs() |> 
  as_tibble() |> 
  ggplot(aes(x = ratio, y = contrast)) +
  geom_vline(xintercept = 1, lty = 11) +
  geom_pointrange(aes(xmin = lower.HPD, xmax = upper.HPD), fill = "black", pch = 21, size = 1) +
  scale_fill_manual(values = websafe_bluepal) +
  scale_x_continuous("Contrast ratio", trans = "log2", limits = c(0.5, 2.1), breaks = c(0.5, 1, 2)) +
  scale_y_discrete(position = "right", labels = c("Shallow -\nLower" ,"Shallow -\nUpper","Upper -\nLower")) +
  theme_bjc() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        axis.title.y = element_blank(),
        plot.title = element_text(size = 10, hjust = 0.5),
        legend.position = "none") 
depth_contr
```

# Combine
```{r fig.height=70/26.5, fig.width=120/25.4}
depth_eff_combined <- 
  cowplot::plot_grid(depth_em, depth_contr, labels = c('a', 'b'), rel_widths = c(1.1, 0.9))
depth_eff_combined
```

# Save
```{r}
ggsave(filename = "output/depth_eff_combined.pdf", plot = depth_eff_combined, height = 70, width = 115, device = cairo_pdf, units = "mm", dpi = 300)
```


# Or if you want density plots
```{r fig.width=55/25.4, fig.height=80/25.4}
depth_dense_plot <- 
  abun_biogeo_model_nodep_final |> 
  emmeans(~  Depth,
           #regrid = "response",  type = "response",
          epred = TRUE) |>                              # At mean distance from GBR) %>% 
  gather_emmeans_draws()  |> 
  ungroup() %>% 
  ggplot(aes(x = .value, y = Depth, fill = Depth)) +
  stat_halfeye(.width = c(0.8, 0.95), point_interval = "median_hdi") +
  theme_bjc() +
  scale_fill_manual(values = websafe_bluepal) +
  #scale_fill_viridis_d(option = "plasma", end = 0.8) +
  scale_y_discrete(position = "right", expand = c(0.1, 0)) +
  scale_x_continuous(limits = c(0,6)) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(colour = "black", size = 10),
        axis.text.y.right = element_text(colour = "black", angle = 270, hjust = 1.2, size = 10),
        axis.text.x = element_text(colour = "black", size = 9),
        strip.text = element_blank()) +
  guides(fill = "none") +
  labs(x = "Mean predators.150m2") +
  facet_wrap(~ Depth, ncol = 1, scales = "free_y")
depth_dense_plot
```



## Trend plots ##

### dGBR * Depth

## Extract range 
```{r}
dGBR_range <- predsum %>% 
  with(list(dGBR = seq_range(dGBR, n = 100)))
```

## Plot
```{r fig.height=80/26.5, fig.width=100/25.4}
abun_dGBR_trend <- 
  abun_biogeo_model_nodep_final |> 
  emmeans(~ dGBR | Depth, var = "dGBR",
           #regrid = "response",  type = "response",
          epred = TRUE,
           at = dGBR_range) %>% 
  gather_emmeans_draws() %>% 
  as_tibble() %>% 
  ggplot(aes(x = dGBR, y = .value)) +
  #stat_lineribbon(colour = "black", fill = "#ccccff", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.6) +
  #stat_lineribbon(colour = "black", fill = "#000099", size = 0.5, point_interval = "median_hdci", .width = 0.85, alpha = 0.9) +
  stat_lineribbon(aes(fill = Depth), colour = "black",  size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.6) +
  stat_lineribbon(aes(fill = Depth), colour = "black", size = 0.5, point_interval = "median_hdci", .width = 0.85, alpha = 0.9) +
  facet_wrap(~Depth, ncol = 1) +
  #scale_colour_viridis_d(option = "plasma", end = 0.8) +
  #scale_fill_viridis_d(option = "plasma", end = 0.8) +
  scale_fill_manual(values = websafe_bluepal) +
  scale_y_continuous(expand = c(0.1, 0)) +
  labs(x = "Dist. GBR shelf (km)", 
       y = "Abundance (No. of predators.150m2)",
       fill = "Credible interval",
       title = "") +
  theme_bjc() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        #strip.text = element_text(colour = "black", size = 9),
        strip.text = element_blank(),
        legend.position = 'none')

abun_dGBR_trend 
```

# Or
```{r fig.height=70/26.5, fig.width=100/25.4}
  abun_biogeo_model_nodep_final |> 
  emmeans(~ dGBR | Depth, var = "dGBR",
           regrid = "response", type = "response",
           at = dGBR_range) %>% 
  as_tibble() %>% 
  ggplot() +
  geom_ribbon(aes(x = dGBR, ymin = lower.HPD, ymax = upper.HPD, fill = Depth)) +
  geom_line(aes(x = dGBR, y = prob, group = Depth), colour = "black") +
  #scale_colour_viridis_d(option = "plasma", end = 0.8) +
  #scale_fill_viridis_d(option = "plasma", end = 0.8) +
  scale_fill_manual(values = websafe_bluepal) +
  #scale_colour_manual(values = c("blue", "dodgerblue4", "dodgerblue")) +
  scale_y_continuous(limits = c(0,18)) +
  labs(x = "Dist. GBR shelf (km)", 
       y = "Abundance (No. of predators.150m2)") +
  theme_bjc() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        strip.text = element_text(colour = "black", size = 9),
        legend.position = 'none') +
  facet_wrap(~Depth, nrow = 1)
```



## dCT * Depth

## Extract range
```{r}
dCT_range <- predsum %>% 
  with(list(dCT = seq_range(dCT, n = 100)))
```

## Plot
```{r fig.height=80/26.5, fig.width=100/25.4}
abun_dCT_trend <- 
  abun_biogeo_model_nodep_final |> 
  emmeans(~ dCT | Depth, var = "dCT",
           #regrid = "response",  type = "response",
          epred = TRUE,
           at = dCT_range) %>% 
  gather_emmeans_draws() %>% 
  as_tibble() %>% 
  ggplot(aes(x = dCT, y = .value)) +
  #stat_lineribbon(colour = "black", fill = "#ccccff", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.6) +
  #stat_lineribbon(colour = "black", fill = "#000099", size = 0.5, point_interval = "median_hdci", .width = 0.85, alpha = 0.9) +
  stat_lineribbon(aes(fill = Depth), colour = "black",  size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.6) +
  stat_lineribbon(aes(fill = Depth), colour = "black", size = 0.5, point_interval = "median_hdci", .width = 0.85, alpha = 0.9) +
  facet_wrap(~Depth, ncol = 1) +
  scale_fill_manual(values = websafe_bluepal) +
  scale_y_continuous(expand = c(0.1, 0)) +
  labs(x = "Dist. Coral Triangle (km)", 
       y = "Abundance (No. of predators.150m2)",
       title = "") +
  theme_bjc() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        #strip.text = element_text(colour = "black", size = 9),
        strip.text = element_blank(),
        legend.position = 'none')

abun_dCT_trend
```



```{r fig.width=165/25.4, fig.height=90/25.4}
abun_biogeo_plot <- 
  cowplot::plot_grid(abun_dCT_trend, abun_dGBR_trend, depth_dense_plot, nrow = 1, labels = c("a", "b", "c"), rel_widths = c(1.1,1,1), label_size = 8, align = "h") +
  theme(plot.background = element_rect(color = "black"))
abun_biogeo_plot
```



# Save
```{r}
ggsave(filename = "output/abun_biogeo_fig.pdf", plot = abun_biogeo_plot, height = 80, width = 160, device = cairo_pdf, units = "mm", dpi = 300)
```










## JUNK ##



# dCT * Depth
```{r fig.height=160/26.5, fig.width=100/25.4}
biogeo_model |> 
  emmeans(~ Depth|dCT, at = biogeogrid, type = 'response',  response = "regrid") |> 
  regrid()  |> 
  as_tibble() |> 
  ggplot() +
  geom_line(aes)
  geom_pointrange(aes(x = Depth, y = prob, ymin = lower.HPD, ymax = upper.HPD, colour = Depth, fill = Depth), size = 1) +
  scale_fill_manual(values = c("#006ad1", "#1e90ff", "#6bb6ff")) +
  scale_colour_manual(values = c("#006ad1", "#1e90ff", "#6bb6ff")) +
  theme_bjc() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        strip.text = element_text(colour = "black", size = 9),
        plot.title = element_text(size = 10, hjust = 0.5),
        legend.position = "none") +
  facet_wrap(~dCT, nrow = 3, strip.position = "right")
```

