---
title: "Abundance modelling - results"
author: "Ben Cresswell"
date: today
date-format: "DD/MM/YYYY"
format: 
  html:
    ## Format
    theme: spacelab
    html-math-method: mathjax
    ## Table of contents
    toc: true
    toc-float: true
    ## Numbering
    number-sections: true
    number-depth: 3
    ## Layout
    fig-caption-location: "bottom"
    fig-align: "center"
    fig-width: 3
    fig-height: 2
    fig-dpi: 100
    tbl-cap-location: top
    ## Code
    code-fold: false
    code-tools: true
    code-summary: "Show the code"
    code-line-numbers: true
    code-block-border-left: "#ccc"
    highlight-style: zenburn
    ## Execution
    execute:
      echo: true
      cache: true
    ## Rendering
    embed-resources: true
documentclass: article
fontsize: 12pt
execute-dir: project
mainfont: Arial
classoption: a4paper
---


```{r}
# Preliminaries
#| label: setup
#| include: false
knitr::opts_chunk$set(cache.lazy = FALSE,
                      tidy = "styler")
options(tinytex.engine = "xelatex")
```


```{r}
#rm(list=ls())

# Load package and functions needed
source("code/packages.R")
source("code/extras.R")
source("code/helper_functions.R")

# Load dataframe and model
#sourceRmd("code/abundance/diw_abund.Rmd")
#sourceRmd("code/abundance/abundance_model_fit.Rmd")

# Or
load("data/Rdata/predsum.Rdata")
load("data/Rdata/abun_biogeo_model.Rda")
```



## Conditional effects plot again to aid interpretation
```{r}
abun_biogeo_model |> conditional_effects() |> plot(ask = FALSE, points = FALSE) # Without raw data
#abun_biogeo_model_nodep_final |> conditional_effects() |> plot(ask = FALSE, points = TRUE) # With raw data overplotted
```

## Summary table

# tidyMCMC is on link scale)
```{r}
# abun_biogeo_model$fit |> 
#   tidyMCMC(estimate.method = 'median',
#              conf.int = TRUE,  conf.method = 'HPDinterval',
#              rhat = TRUE, ess = TRUE)
```


```{r}
abun_biogeo_model |> 
  as_draws_df() |>    # Extract draws
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  rhat, ess_bulk, ess_tail,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1))|> 
  as_tibble() |> 
  mutate(median = round(median, 3),
         lower = round(lower, 3),
         upper = round(upper, 3))
```



Need something that is interpretable




```{r fig.width=50/25.4, fig.height=100/25.4}
abun_effects_draws <- 
abun_biogeo_model |> 
 gather_draws(`^b_.*`, regex=TRUE)  |> 
  filter(!grepl("Intercept", .variable))   |> 
  mutate(Variable = factor(case_when(grepl("CT", .variable) ~ "Distance\nCT (km)",
                                     grepl("GBR", .variable) ~ "Distance to\n GBR (km)",
                                     grepl("Reef_100", .variable) ~ "Reef w/in\n100km (kha)",
                                     grepl("Reef_1_Area", .variable) ~ "Reef size\n(kha)")))
```




```{r fig.width=50/25.4, fig.height=100/25.4}
#abun_effects_plot <- 
  abun_effects_draws |> 
  ggplot(aes(x = .value, y = Variable, fill = Variable, group = Variable)) +
  geom_vline(xintercept = 0, lty = 11) +
  stat_halfeye(.width = c(0.8, 0.95), point_interval = "median_hdi", slab_colour = "black", slab_linewidth = 0.5) +
  theme_bjc() +
  scale_fill_manual(values = cb_palette) +
  scale_y_discrete(position = "left", expand = c(0.05, 0)) +
  #annotate(geom = "text", label = "P=0.53", x = -0.4, y = 1, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5), limits = c(-0.9, 0.6)) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(colour = "black", size = 9),
        #axis.text.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 8),
        strip.text = element_blank()) +
  guides(fill = "none") +
  labs(x = "Scaled marginal effect")
abun_effects_plot
```

```{r}
ggsave(filename = "output/abun_effects_fig.pdf", plot = abun_effects_plot, height = 80, width = 50, device = cairo_pdf, units = "mm", dpi = 300)
```


## Want to get the credible intervals showing up on the actual slabs so need to do these manually:

### dGBR
```{r fig.width=50/25.4, fig.height=25/25.4}
dGBR_abun_dens <- 
abun_effects_draws |> 
  filter(grepl("GBR", .variable)) |> 
  ggplot(aes(x = .value, y = Variable, fill = Variable, group = Variable)) +
  geom_vline(xintercept = 0, lty = 11) +
  stat_halfeye(aes(fill = after_stat(ggdist::cut_cdf_qi(cdf,
                                         .width = c(0.8,0.95,1),
                                         labels = scales::percent_format()))),
    .width = c(0.8, 0.95), point_interval = "median_qi", slab_colour = "black", slab_linewidth = 0.5,
    point_size = 2) +
  theme_bjc() +
  scale_fill_manual(values = dGBR_pal) +
  scale_y_discrete(position = "left", expand = c(0.07, 0)) +
  #annotate(geom = "text", label = "P=0.53", x = -0.4, y = 1, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5), limits = c(-0.85, 0.55), minor_breaks = NULL) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        plot.margin = unit(c(0,0,0,0), "mm")) +
  guides(fill = "none") +
  labs(x = "Scaled marginal effect")

dGBR_abun_dens
```


### dCT
```{r fig.width=50/25.4, fig.height=25/25.4}
dCT_abun_dens <- 
abun_effects_draws |> 
  filter(grepl("CT", .variable)) |> 
  ggplot(aes(x = .value, y = Variable, fill = Variable, group = Variable)) +
  geom_vline(xintercept = 0, lty = 11) +
  stat_halfeye(aes(fill = after_stat(ggdist::cut_cdf_qi(cdf,
                                         .width = c(0.8,0.95,1),
                                         labels = scales::percent_format()))),
    .width = c(0.8, 0.95), point_interval = "median_qi", slab_colour = "black", slab_linewidth = 0.5,
    point_size = 2) +
  theme_bjc() +
  scale_fill_manual(values = dCT_pal) +
  scale_y_discrete(position = "left", expand = c(0.07, 0)) +
  #annotate(geom = "text", label = "P=0.53", x = -0.4, y = 1, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5), limits = c(-0.85, 0.55), minor_breaks = NULL) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        plot.margin = unit(c(0,0,0,0), "mm")) +
  guides(fill = "none") +
  labs(x = "Scaled marginal effect")

dCT_abun_dens
```


### Reef_1_Area
```{r fig.width=50/25.4, fig.height=25/25.4}
Reef1Area_abun_dens <- 
abun_effects_draws |> 
  filter(grepl("Area", .variable)) |> 
  ggplot(aes(x = .value, y = Variable, fill = Variable, group = Variable)) +
  geom_vline(xintercept = 0, lty = 11) +
  stat_halfeye(aes(fill = after_stat(ggdist::cut_cdf_qi(cdf,
                                         .width = c(0.8,0.95,1),
                                         labels = scales::percent_format()))),
    .width = c(0.8, 0.95), point_interval = "median_qi", slab_colour = "black", slab_linewidth = 0.5,
    point_size = 2) +
  theme_bjc() +
  scale_fill_manual(values = Reef_1_Area_pal) +
  scale_y_discrete(position = "left", expand = c(0.07, 0)) +
  #annotate(geom = "text", label = "P=0.53", x = -0.4, y = 1, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5), limits = c(-0.85, 0.55), minor_breaks = NULL) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        plot.margin = unit(c(0,0,0,0), "mm")) +
  guides(fill = "none") +
  labs(x = "Scaled marginal effect")

Reef1Area_abun_dens
```


### Reef_100
```{r fig.width=50/25.4, fig.height=25/25.4}
Reef100_abun_dens <- 
abun_effects_draws |> 
  filter(grepl("100", .variable)) |> 
  ggplot(aes(x = .value, y = Variable, fill = Variable, group = Variable)) +
  geom_vline(xintercept = 0, lty = 11) +
  stat_halfeye(aes(fill = after_stat(ggdist::cut_cdf_qi(cdf,
                                         .width = c(0.8,0.95,1),
                                         labels = scales::percent_format()))),
    .width = c(0.8, 0.95), point_interval = "median_qi", slab_colour = "black", slab_linewidth = 0.5,
    point_size = 2) +
  theme_bjc() +
  scale_fill_manual(values = Reef_100_pal) +
  scale_y_discrete(position = "left", expand = c(0.08, 0)) +
  #annotate(geom = "text", label = "P=0.53", x = -0.4, y = 1, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5), limits = c(-0.85, 0.55), minor_breaks = NULL) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(colour = "black", size = 7),
#        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 7),
#        axis.text.x = element_blank(),
        plot.margin = unit(c(0,0,2,0), "mm")) +
  guides(fill = "none") +
  labs(x = "Scaled marginal effect")

Reef100_abun_dens
```

# Combine density plots for LHS of main figure
```{r fig.width=50/25.4, fig.height=100/25.4}
abun_plot_LHS <- 
dGBR_abun_dens /
dCT_abun_dens /
Reef1Area_abun_dens /
Reef100_abun_dens 



abun_plot_LHS <- plot_grid(dGBR_abun_dens, dCT_abun_dens, Reef1Area_abun_dens, Reef100_abun_dens, ncol = 1) +
    theme(plot.background = element_rect(fill = "white", colour = "black"))




abun_plot_LHS <- abun_plot_LHS +
  theme(plot.background = element_rect(fill = "white", colour = "black"))

abun_plot_LHS
```




# Now can do trend plots to go with these:

## dGBR

```{r fig.height=25/26.5, fig.width=50/25.4}
dGBR_abun_trend <- 
abun_biogeo_model |> 
  emmeans(~ dGBR, type = 'link', 
          epred = TRUE, 
          at = predsum |> with(list(dGBR = seq_range(dGBR, n = 100)))) |>  # Just building new data grid within emmeans
  regrid()  |> 
  gather_emmeans_draws()  |> 
  as_tibble() |> 
  ggplot(aes(x = dGBR, y = .value)) +
  geom_line(aes(group = .draw), colour = "#999933", alpha = 0.1) +
  stat_lineribbon(colour = "black", size = 0.4, point_interval = "median_qi", .width = 0.0001, alpha = 1) +
  theme_bjc() +
  scale_x_continuous("Distance GBR (km)", minor_breaks = NULL) +
  scale_y_continuous("No. preds (.150m2)", 
                     expand = c(0, 0.1),
                     limits = c(0,10),
                     breaks = c(0, 5, 10),
                     position = "right",
                     minor_breaks = NULL) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(colour = "black", size = 7),
        axis.title.x = element_text(colour = "black", size = 7),
        axis.text.x = element_text(colour = "black", size = 7),
        #plot.margin = unit(c(0,0,2,0), "mm"),
        plot.background = element_rect(fill = "white", colour = "black"),
        legend.position = "none")

dGBR_abun_trend
```

# dCT
```{r fig.height=25/26.5, fig.width=50/25.4}
dCT_abun_trend <- 
abun_biogeo_model |> 
  emmeans(~ dCT, type = 'link', 
          epred = TRUE, 
          at = predsum |> with(list(dCT = seq_range(dCT, n = 100)))) |>  # Just building new data grid within emmeans
  regrid()  |> 
  gather_emmeans_draws()  |> 
  as_tibble() |> 
  ggplot(aes(x = dCT, y = .value)) +
  geom_line(aes(group = .draw), colour = "#44aa99", alpha = 0.1) +
  stat_lineribbon(colour = "black", size = 0.4, point_interval = "median_hdci", .width = 0.0001, alpha = 1) +
  theme_bjc() +
  scale_x_continuous("Distance CT (km)") +
  scale_y_continuous("No. preds (.150m2)", 
                     expand = c(0, 0.1),
                     limits = c(0,10),
                     breaks = c(0, 5, 10),
                     position = "right",
                     minor_breaks = NULL) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(colour = "black", size = 7),
        axis.title.x = element_text(colour = "black", size = 7),
        axis.text.x = element_text(colour = "black", size = 7),
        plot.background = element_rect(fill = "white", colour = "black"),
        #plot.margin = unit(c(0,0,2,0), "mm"),
        legend.position = "none")


dCT_abun_trend
```

# Reef_1_Area
```{r fig.height=25/26.5, fig.width=50/25.4}
Reef1Area_abun_trend <- 
abun_biogeo_model |> 
  emmeans(~ Reef_1_Area, type = 'link', 
          epred = TRUE, 
          at = predsum |> with(list(Reef_1_Area = seq_range(Reef_1_Area, n = 100)))) |>  #
  regrid()  |> 
  gather_emmeans_draws()  |> 
  as_tibble() |> 
  ggplot(aes(x = Reef_1_Area, y = .value)) +
  geom_line(aes(group = .draw), colour = "#aa4499", alpha = 0.1) +
  stat_lineribbon(colour = "black", size = 0.4, point_interval = "median_hdci", .width = 0.0001, alpha = 1) +
  theme_bjc() +
  scale_x_continuous("Reef size (kha)") +
  scale_y_continuous("No. preds (.150m2)", 
                     expand = c(0, 0.1),
                     limits = c(0,10),
                     breaks = c(0, 5, 10),
                     position = "right",
                     minor_breaks = NULL) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(colour = "black", size = 7),
        axis.title.x = element_text(colour = "black", size = 7),
        axis.text.x = element_text(colour = "black", size = 7),
        #plot.margin = unit(c(0,0,2,0), "mm"),
        plot.background = element_rect(fill = "white", colour = "black"),
        legend.position = "none")


Reef1Area_abun_trend
```


# Reef_100
```{r fig.height=25/26.5, fig.width=50/25.4}
Reef100_abun_trend <- 
abun_biogeo_model |> 
  emmeans(~ Reef_100, type = 'link', 
          epred = TRUE, 
          at = predsum |> with(list(Reef_100 = seq_range(Reef_100, n = 100)))) |>  #
  regrid()  |> 
  gather_emmeans_draws()  |> 
  as_tibble() |> 
  ggplot(aes(x = Reef_100, y = .value)) +
  geom_line(aes(group = .draw), colour = "#332288", alpha = 0.1) +
  stat_lineribbon(colour = "black", size = 0.4, point_interval = "median_hdci", .width = 0.0001, alpha = 1) +
  theme_bjc() +
  scale_x_continuous("Reef isol (kha reef w/in 100km)") +
  scale_y_continuous("No. preds (.150m2)", 
                     expand = c(0, 0.1),
                     limits = c(0,12),
                     breaks = c(0, 5, 10),
                     position = "right",
                     minor_breaks = NULL) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(colour = "black", size = 7),
        axis.title.x = element_text(colour = "black", size = 7),
        axis.text.x = element_text(colour = "black", size = 7),
        #plot.margin = unit(c(0,0,2,0), "mm"),
        plot.background = element_rect(fill = "white", colour = "black"),
        legend.position = "none")


Reef100_abun_trend
```




# Combine to make RHS of plot
```{r fig.height=100/26.5, fig.width=50/25.4}
abun_plot_RHS <- 
dGBR_abun_trend /
dCT_abun_trend /
Reef1Area_abun_trend /
Reef100_abun_trend +
  ylab("No. preds (.150m2)")

abun_plot_RHS


```




```{r fig.height=100/26.5, fig.width=50/25.4}
abun_plot_RHS2 <- plot_grid(dGBR_abun_trend, dCT_abun_trend, Reef1Area_abun_trend, Reef100_abun_trend, ncol = 1) +
    theme(plot.background = element_rect(fill = "white", colour = "black")) +
  ylab("No. preds (.150m2)")

abun_plot_RHS2
```

```{r}
ggsave(filename = "output/abun_trends_fig.pdf", plot = abun_plot_RHS, height = 100, width = 50, device = cairo_pdf, units = "mm", dpi = 300)
```



```{r fig.height=100/26.5, fig.width=100/25.4}
abun_plot_final <- 
  plot_grid(abun_plot_LHS, abun_plot_RHS, ncol = 2, rel_widths = c(0.5,1)) +
  theme(plot.background = element_rect(fill = "white", colour = "black"),
        plot.margin = unit(c(0,10,0,0), "mm"))


abun_plot_final2 <- 
  abun_plot_LHS + 
  abun_plot_RHS +
  plot_layout(ncol = 2, widths = c(0.5,1)) +
  plot_annotation(tag_levels = 'A') +
  theme(plot.background = element_rect(fill = "white", colour = "black"),
        plot.margin = unit(c(0,10,0,0), "mm"))
```


```{r}
abun_plot_final
abun_plot_final2

```

# Save
```{r}
ggsave(filename = "output/abun_effects_fig.pdf", plot = abun_plot_final, height = 100, width = 100, device = cairo_pdf, units = "mm", dpi = 300)
```



