---
title: "Abundance ~ biogeographic model fit"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    #fig_height: 
    fig_width: 140/25.4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```



# Load predsum data
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
#sourceRmd("code/abundance/diw_abund.Rmd")
load("data/Rdata/predsum.Rdata")
```

# Check
```{r}
head(predsum)
```


# Introduction
Predator abundance is modeled against the main isolation-type metrics in a separate document (abun_isol_brm.Rmd). We have a number of other variables that we might expect to explain some of the other variability in abundance. These are mainly geographic and candidate variables are as follows:

-------------         ------------------------------------------------------------------------------
## DEPTH ##
**Depth_m**           Depth as continuous variable
**Depth_bin**         Ordered 10 level variable - 10m depth bins
**Depth_bin_meso**    3 factor categorical variable - depth bins by photic zone. NOTE: This factor has now ALSO been recoded as **'Depth'**

## BIOGEOGRAPHIC ##
**Site_lat**          Continuous variable - distance from equator in ° latitude - not very useful. Could instead use:
**dCT**               Continuous variable - distance from Coral Triangle in km. More usable.
**Site_long**         Continuous variable - distance from UTC in ° long - also not very useful. Could instead use:
**dGBR**              Continuous variable - distance from GBR shelf in km. More usable.
**Region**            4 factor categorical variable - Coral Sea region, Far North, North, Central, South. Could be used instead of any of the above

-------------     ------------------------------------------------------------------------------



## Of these we'll start with:
- Depth
- dCT
- dGBR

# Check multicolinearity and distribution of predictors
```{r message=FALSE, warning=FALSE, include=FALSE}
car::scatterplotMatrix(~ No_Preds + dCT + dGBR + Depth_bin + Depth, data = predsum, diagonal = list(method = 'boxplot'))
```

```{r}
predsum |> 
  ggplot() +
  #geom_histogram(aes(x = dGBR))
  geom_histogram(aes(x = dCT))

```



# Set model formula
```{r}
abun_biogeo_form <- bf(No_Preds ~ Depth*scale(dCT) + Depth*scale(dGBR) + (1|Site), zi ~ 1, family = zero_inflated_negbinomial(link = "log"))
```

## Check what priors we need to estimate
```{r}
get_prior(abun_biogeo_form, data = predsum)
```


## Set priors (start with the same as the isol model)
```{r}
abun_biogeo_priors <-  
  prior(normal(1.5, 2), class = 'Intercept') +
  prior(normal(0, 0.5), class = 'b') +
  prior(gamma(2, 2), class = 'shape') +
  prior(logistic(0, 1), class = 'Intercept', dpar = "zi")
```


## Fit model: 
```{r}
abun_biogeo_model_priorsonly <- 
  brm(abun_biogeo_form,
  data = predsum,
  prior = abun_biogeo_priors,
  sample_prior = 'only',              
  chains = 3,
  cores = 3,
  iter = 5000, 
  warmup = 1000, 
  thin = 4,
  seed = 1234,
  silent = 0,
  refresh = 0,
  backend = "cmdstanr")
```


## Check how wide priors are
```{r fig.width=140/25.4}
abun_biogeo_model_priorsonly |> conditional_effects() |> 
   plot(points = TRUE, ask = FALSE, plot = TRUE) 
```


## Fit with data
```{r cache=TRUE}
abun_biogeo_model <- abun_biogeo_model_priorsonly |> update(sample_prior = "yes", refresh = 0)
```

## Check influence of priors
```{r}
abun_biogeo_model |> SUYR_prior_and_posterior()
```


# Model validation

## DHARMa residuals
```{r}
resids <- make_brms_dharma_res(abun_biogeo_model, integerResponse = FALSE)
testUniformity(resids)
plotResiduals(resids, form = factor(rep(1, nrow(predsum))))
plotResiduals(resids, quantreg = TRUE)
testDispersion(resids) # Dispersion plot (dispersion = variance/mean) - we assume mean and variance are the same, so we need to check
```

## Residuals look good. 

## Chain checks

```{r}
abun_biogeo_model |> plot(ask = FALSE)
# Trace plot
mcmc_plot(abun_biogeo_model, type='trace')
# Autocorrelation
mcmc_plot(abun_biogeo_model, type='acf_bar')
# Rhat - how well the chains converged
mcmc_plot(abun_biogeo_model, type='rhat_hist')
# Percentage effectives - needs to be greater than 0.5
mcmc_plot(abun_biogeo_model, type='neff_hist')
```



## PP checks
```{r}
# Logged
pred <- posterior_predict(abun_biogeo_model)
bayesplot::ppc_dens_overlay(y = log(predsum$No_Preds), 
                            yrep = log(pred[1:100,]))
```
Posterior draws fit the data nicely:
y = the observed data
yrep = the predictive(posterior) distribution


### NEED TO DO THE ZI WORK HERE ###
```{r}

```

# Model investigation 
## Conditional effects plot again to aid interpretation
```{r}
abun_biogeo_model |> conditional_effects() |> plot(ask = FALSE, points = FALSE) # Without raw data
#abun_biogeo_model |> conditional_effects() |> plot(ask = FALSE, points = TRUE) # With raw data overplotted
```

## Summary table
```{r}
abun_biogeo_model |> 
  as_draws_df() |>    # Extract draws
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  rhat, ess_bulk, ess_tail,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1))|> 
  as_tibble() |> 
  mutate(median = round(median, 3),
         lower = round(lower, 3),
         upper = round(upper, 3))
```



# Try with region as combined predictor of horizontal position

# Priors should be good but we'll need to specify the formula

```{r}
region_form <- bf(No_Preds ~ Depth*Region + (1|Site), zi ~ 1, family = zero_inflated_negbinomial(link = "log"))
```

## Fit with data
```{r cache=TRUE}
abun_biogeo_model2 <- abun_biogeo_model |> update(formula = region_form, newdata = predsum, sample_prior = "yes", refresh = 0)
```

## Quick chain check
```{r}
abun_biogeo_model2 |> plot(ask = FALSE)
```

## DHARMa residuals
```{r}
resids2 <- make_brms_dharma_res(abun_biogeo_model2, integerResponse = FALSE)
testUniformity(resids2)
plotResiduals(resids2, form = factor(rep(1, nrow(predsum))))
plotResiduals(resids2, quantreg = TRUE)
testDispersion(resids2) # Dispersion plot (dispersion = variance/mean) - we assume mean and variance are the same, so we need to check
```

# R2
```{r}
abun_biogeo_model |> bayes_R2()
abun_biogeo_model2 |> bayes_R2()
```



```{r}
loo_compare(brms::loo(abun_biogeo_model), brms::loo(abun_biogeo_model2))
```

            elpd_diff se_diff
abun_biogeo_model   0.0       0.0   
abun_biogeo_model2 -3.9       2.8

The full continuous predictor model comes out best. 

## Assign that as final model 
```{r}
abun_biogeo_model_final <- abun_biogeo_model
```

## Save
```{r}
save(abun_biogeo_model_final, file = "data/Rdata/abun_biogeo_model_final.Rda")
```








