---
title: "Coral Sea ROV Fish dataPredator abundance wrangling - "
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Load predator data, packages
```{r message=FALSE, warning=FALSE, include=FALSE, fig.width = 150/25.4, fig.height = 90/25.4}
source("code/packages.R")
source("code/wrangling/diw_rawdata.R")
# Or
load('data/Rdata/preds.Rdata')
# Check data
preds
# Distribution of response variable
preds %>% ggplot() +  
  geom_histogram(aes(x=Number), binwidth = 1, color="black", fill="grey") 
  scale_x_continuous("No. of predators")
```



## Lots of zeros or near-zeros and we have "outliers" in the high numbers - 40+

## Let's check what our ranges are
```{r}
preds %>% 
  summarise_at(vars(Number), list(median = median, mean = mean, SD = sd, SE = std.error,
                                  CI5 = ~quantile(., probs = 0.05), CI95 = ~quantile(., probs = 0.95), 
                                  CI1 = ~quantile(., probs = 0.01), CI99 = ~quantile(., probs = 0.99), 
                                  Q1 = ~quantile(., probs = 0.25), Q3=~quantile(., probs = 0.75)))
```

So the average number of preds observed per transect is around 1 (median and mean) with a fair bit of variability (SD is nearly double the mean)
99% of observations are 14 or less, so we'll round down these outliers to minimise their effect. 


## Next we'll see how many observations this will affect:
```{r}
preds %>% filter(Number >14)
```
  

12 observations - less than 1% of all the observations in the dataset. We can leave in, remove or round down


## Summarise predator counts by T_ID (include all other vars of interest here)

Can sort the outliers here if required

```{r}
predsum <- 
   preds %>% 
   #filter(grepl('out', Habitat)) %>% droplevels() %>% 
   #mutate(across(Number, ~  ifelse(. > 14, 14, .))) %>% # Round down outliers if required
   group_by(Reef_1, Reef_2, Site, Site_Depth,                        # Grouping variables of reef and site - only going to need Reef_1 (overarching reef system)
            Site_lat, Site_long,                         # Geographic variables. Don't think these will be that useful in a model (meaningless units) but could use...
            dCT, dGBR, Region,                           # Distance from coral triangle, GBR and region (categorical) are all possible proxies for lat/long
            Depth, Depth_m, Depth_bin, Depth_bin_meso,   # Can model depth as continuous or as a binned variable  
            Reef_1_Area, Reef_size,                      # Same with reef area - as continuous or binned
            Isol, IsolRank, Reef_100,                    # Isolation score is a continuous variable, rank is ordered, km2 of reef within 100km
            Aspect_descriptive, Aspect_scale,             # Both of these are actually ordered categorical despite one appearing to be numeric
            Temperature, Benthic_Complexity,             #
            Situation, Habitat, Zone,
            T_ID) %>%
   summarise(No_Preds =sum(Number)) %>%  # All good - 514
   ungroup()
```


```{r}
predsum %>% str()
```

# How many predators in total across study?
```{r}
predsum %>% 
  summarise(Total_preds = sum(No_Preds))
```

# No of families and species 


# Save if required
```{r}
#save(predsum, file = 'data/Rdata/predsum.Rdata')
#load(file = 'data/Rdata/predsum.Rdata')
```


## Predator abundance by reef

### Not expecting any patterns here but worth a quick check
```{r fig.width = 390/25.4, fig.height = 100/25.4}
predsum %>% 
  group_by(Reef_1) %>% 
  summarise_at(vars(No_Preds), list(median = median, mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() +
  geom_pointrange(aes(x = Reef_1, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```
Ordered by most to least isolated

## Effect of depth?

### As continuous
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% 
  ggplot() + 
  geom_point(aes(x = Depth_m, y = No_Preds)) +
  #geom_smooth(aes(x = Depth_m, y = No_Preds), method = "glm") +
  geom_smooth(aes(x = Depth_m, y = No_Preds), method = "gam") +
  theme_minimal() +
  ylab("No of predators per transect") +
  xlab("Depth (m)") +
  scale_y_continuous(trans = 'log2')
```


Too much noise....



### 10m depth bins
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% 
  group_by(Depth_bin) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Depth_bin, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Depth_bin, y = mean)) +
  geom_line(aes(x = Depth_bin, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Depth (10m bins)")
```



### Photic zone bins
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% 
  group_by(Depth_bin_meso) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Depth_bin_meso, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Depth_bin_meso, y = mean)) +
  geom_line(aes(x = Depth_bin_meso, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Depth zone")
```




# Region (proxy for latitude)
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% 
  group_by(Region) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Region, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Region, y = mean)) +
  geom_line(aes(x = Region, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Mean no of predators per transect ± SE)") +
  xlab("Region)")

```


# Region * Depth
```{r fig.width = 170/25.4, fig.height = 90/25.4}
predsum %>% 
  group_by(Depth_bin, Region) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Depth_bin, y = mean, ymin = mean-SE, ymax = mean+SE, colour = Region)) +
  geom_point(aes(x = Depth_bin, y = mean, colour = Region)) +
  geom_line(aes(x = Depth_bin, y = mean, group = Region, colour = Region)) +
  theme_minimal() +
  ylab("Mean no of predators per transect ± SE)") +
  xlab("Depth (10m bins)") +
  facet_wrap(~ Region)
```


# Region * Depth bin meso
```{r fig.width = 170/25.4, fig.height = 90/25.4}
predsum %>% 
  group_by(Depth_bin_meso, Region) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Depth_bin_meso, y = mean, ymin = mean-SE, ymax = mean+SE, colour = Region)) +
  geom_point(aes(x = Depth_bin_meso, y = mean, colour = Region)) +
  geom_line(aes(x = Depth_bin_meso, y = mean, group = Region, colour = Region)) +
  theme_minimal() +
  ylab("Mean no of predators per transect ± SE)") +
  xlab("Depth bin") +
  facet_wrap(~ Region)
```



# Or...
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% 
  mutate(Depth_bin = fct_relevel(Depth_bin, rev)) %>% 
  group_by(Depth_bin, Region) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_point(aes(x = Region, y = Depth_bin, colour = mean+(SE*2), fill = mean+(SE*2), group = mean+(SE*2), size = mean+(SE*2)), position = position_jitter(width = 0.3, height = 0, seed = 1)) +
  geom_point(aes(x = Region, y = Depth_bin, size = mean, group = 1), colour = "black", fill = "white", pch = 21, position = position_jitter(width = 0.3, height = 0, seed = 1)) +
  theme_minimal() +
  scale_radius(range = c(0, 15), name = "Abundance") +
  scale_fill_gradient(guide = "legend",low = "white", high = "dodgerblue", name = "Abundance") +
  scale_colour_gradient(guide = "legend", low = "white", high = "dodgerblue", name = "Abundance") +
  scale_y_discrete(expand = c(0,1)) +
  ylab("Depth (10m bins)") +
  xlab("Region")
```


```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% 
  mutate(Depth_bin_meso = fct_relevel(Depth_bin_meso, rev)) %>% 
  group_by(Depth_bin_meso, Region) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_point(aes(x = Region, y = Depth_bin_meso, colour = mean+(SE*2), fill = mean+(SE*2), group = mean+(SE*2), size = mean+(SE*2))) +
  geom_point(aes(x = Region, y = Depth_bin_meso, size = mean, group = 1), colour = "black", fill = "white", pch = 21) +
  theme_minimal() +
  scale_radius(range = c(0, 15), name = "Abundance") +
  scale_fill_gradient(guide = "legend",low = "white", high = "dodgerblue", name = "Abundance") +
  scale_colour_gradient(guide = "legend", low = "white", high = "dodgerblue", name = "Abundance") +
  scale_y_discrete(expand = c(0,1)) +
  ylab("Depth bin") +
  xlab("Region")
```




## EFfect of latitude?
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% ggplot() + 
  geom_point(aes(x = Site_lat, y = No_Preds))  +
  geom_smooth(aes(x = Site_lat, y = No_Preds), method = "glm") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 12))
```


# Could also use distance from Coral Triangle - presumably will marry up with general latitude patterns (although it's mitigated by offshore reefs being near to Solomon Isl which is technically CT)
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% ggplot() + 
  geom_point(aes(x = dCT, y = No_Preds)) +
  geom_smooth(aes(x = dCT, y = No_Preds), method = "glm") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 12))

```



## Other spatial drivers?

### Can deal with longitude and distance from GBR at the same time
```{r fig.width = 130/25.4, fig.height = 70/25.4}
# Longitude
predsum %>% ggplot() + 
  geom_point(aes(x = Site_long, y = No_Preds)) +
  geom_smooth(aes(x = Site_long, y = No_Preds), method = "glm") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 12))

# Distance from GBR
predsum %>% ggplot() + 
  geom_point(aes(x = dGBR, y = No_Preds)) +
  geom_smooth(aes(x = dGBR, y = No_Preds), method = "glm") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 12))

```
  

### Looks like some attenuation with distance offshore?



## Benthic complexity?
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% 
  filter(Benthic_Complexity < 30) %>% 
  mutate(Benthic_Complexity = factor(Benthic_Complexity, ordered = TRUE)) %>% 
  group_by(Benthic_Complexity) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Benthic_Complexity, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Benthic_Complexity, y = mean)) +
  geom_line(aes(x = Benthic_Complexity, y = mean, group = 1)) +
  geom_smooth(aes(x = Benthic_Complexity, y = mean, group = 1), method = "loess") +
  theme_minimal() +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Benthic complexity score")
```

## Aspect?
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% 
  mutate(Aspect = factor(Aspect_scale, ordered = TRUE)) %>% 
  group_by(Aspect) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Aspect, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Aspect, y = mean)) +
  geom_line(aes(x = Aspect, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Aspect score (0-5)")
```
# OR

```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% 
  group_by(Aspect_descriptive) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() + 
  geom_pointrange(aes(x = Aspect_descriptive, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Aspect_descriptive, y = mean)) +
  geom_line(aes(x = Aspect_descriptive, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Aspect type")
```


# Temp?
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% ggplot() + 
  geom_point(aes(x = Temperature, y = No_Preds)) +
  geom_smooth(aes(x = Temperature, y = No_Preds, method = "glm")) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 12))

```



## Isolation

### See how the individual reefs stack up
```{r fig.width = 390/25.4, fig.height = 100/25.4}
predsum %>% 
  select(Reef_1, Isol) %>% 
  distinct() %>%  
  mutate(Reef_1 = fct_reorder(Reef_1, Isol)) %>% 
  ggplot(aes(x = Reef_1, y = Isol)) +
  geom_point()

predsum %>% 
  select(Reef_1, Reef_100) %>% 
  distinct() %>%  
  mutate(Reef_1 = fct_reorder(Reef_1, -Reef_100)) %>% 
  ggplot(aes(x = Reef_1, y = Reef_100)) +
  geom_point()
```




```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum %>% ggplot() +
  #geom_point(aes(x = Isol, y = No_Preds)) +
  geom_smooth(aes(x = Isol, y = No_Preds), method = "glm") +
  theme_minimal() 
  scale_y_continuous(trans = scales::log2_trans())
```

# Could just use Reef_100?
```{r}
predsum %>% ggplot() +
  geom_point(aes(x = Reef_100, y = No_Preds)) +
  geom_smooth(aes(x = Reef_100, y = No_Preds), method = "glm") +
  theme_minimal() +
  scale_y_continuous(trans = scales::log10_trans())
```




# Reef size

```{r}
predsum %>% 
  slice_max(No_Preds, n = 20)
#  summary(No_Preds)
```


```{r}
preds %>% 
  filter(T_ID == "Feb_21-HERA-01-D1-Dive_1-T6")
```


```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum  %>%  
  ggplot() +
  geom_point(aes(x = Reef_1_Area, y = No_Preds)) +
  geom_smooth(aes(x = Reef_1_Area, y = No_Preds), method = "glm") +
  theme_minimal() +
  scale_y_continuous(limits = c(0,12)) +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Reef size")
```


# As categorical
```{r fig.width = 150/25.4, fig.height = 90/25.4}
predsum  %>%  
  group_by(Reef_size) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() +
  geom_pointrange(aes(x = Reef_size, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Reef_size, y = mean)) +
  geom_line(aes(x = Reef_size, y = mean, group = 1)) +
  theme_minimal() +
  ylab("No of preds per transect (± SE)") +
  xlab("Reef size")
```
