---
title: "Coral Sea ROV predator abundance wrangling"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Load predator data, packages
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/wrangling/diw_rawdata.R")
# Or
#load('data/Rdata/preds.Rdata')
# Check data
preds

# Distribution of abundance - per transect
preds |> 
  group_by(T_ID) |> 
  summarise(Transect_Abun = sum(Number)) |> 
  ggplot() +  
  geom_histogram(aes(x=Transect_Abun), binwidth = 1, color="black", fill="grey") 
  scale_x_continuous("No. of predators")
```

Lots of zeros or near-zeros and we have "outliers" in the high numbers - probably caused by schools (e.g. Lutjanus karisma)

## Can check by redoing the histogram - per each observation
```{r}
preds |> 
  ggplot() +  
  geom_histogram(aes(x=Number), binwidth = 1, color="black", fill="grey") 
  scale_x_continuous("No. of predators")

```
So most observations are 1 fish, but looks like there were some schools of 10 or more

We can check this by looking at individual observations of large numbers of fishes, to see if likely to be a user error
First need to check what our ranges are. We know there are lots of zeros but what about at the top end of things..
```{r}
preds |> 
  summarise_at(vars(Number), list(median = median, mean = mean, SD = sd, 
                                  CI95 = ~quantile(., probs = 0.95), 
                                  CI99 = ~quantile(., probs = 0.99)))
```

So the average number of preds observed per transect is around 1 (median and mean) with a fair bit of variability
Only 1% of observations are 14 or more - these should be the schooling or erroneous observations:
```{r}
preds |> filter(Number >14)
```
Yep, these are all schooling types, but only 12 observations - less than 1% of all the observations in the dataset. We can either leave in, remove or round down. Will leave in for now as it's so few observations. The important thing is we weren't documenting schools of things that don't school!


## Create predsum tibble - this will be the main tbl for the abundance analysis
Summarise predator counts by T_ID (include all other vars of interest here)
```{r}
predsum <- 
   preds |> 
   group_by(Reef_1, Reef_2, Site,              # Grouping vars of reef & site
            Site_lat, Site_long,               # Geographic vars
            dCT, dGBR, Region,                 # Other NSEW vars
            Reef_1_Area, Reef_100,             # Isolation vars  
            Depth, Site_Depth,                 # Don't need this here, but do need for rarefaction work
            Aspect_descriptive, Aspect_scale,  # Might not need but putting in for now
            Temperature, Benthic_Complexity,   # Ditto
            Situation, Habitat, Zone,          # Ditto
            T_ID) |> 
   summarise(No_Preds =sum(Number))  |>   # All good - 538 - same as metadata
   ungroup()
```


```{r}
predsum |> str() # Prefer str over glimpse as it gives you number of levels.
#predsum  |> glimpse()
```

# How many predators in total across study?
```{r}
predsum |> 
  summarise(Total_preds = sum(No_Preds))
```
1822 - note this is different from the number of rows in the preds file (1312) as that groups observations of >1 individual into single rows


# No of families and species - can get this directly from the preds file. See diw_rawdata.R

# Save if required
```{r}
#save(predsum, file = 'data/Rdata/predsum.Rdata')
#load(file = 'data/Rdata/predsum.Rdata')
```


## Predator abundance by reef

### Not expecting any patterns here but worth a quick check
```{r fig.width = 390/25.4, fig.height = 100/25.4}
predsum |> 
  group_by(Reef_1) %>% 
  summarise_at(vars(No_Preds), list(median = median, mean = mean, SD = sd, SE = std.error)) %>% 
  ggplot() +
  geom_pointrange(aes(x = Reef_1, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```


# Region (proxy for latitude)
```{r}
predsum |> 
  group_by(Region) |> 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) |> 
  ggplot() + 
  geom_pointrange(aes(x = Region, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Region, y = mean)) +
  geom_line(aes(x = Region, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Mean no of predators per transect ± SE)") +
  xlab("Region)")
```


## EFfect of latitude?
```{r}
predsum |> 
  ggplot() + 
  geom_point(aes(x = -Site_lat, y = No_Preds))  +
  geom_smooth(aes(x = -Site_lat, y = No_Preds), method = "glm") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 12)) +
  scale_x_continuous(name = "Latitude (°S)")
```


# Could also use distance from Coral Triangle - presumably will marry up with general latitude patterns (although it's mitigated by offshore reefs being near to Solomon Isl which is technically CT)
```{r}
predsum |> 
  ggplot() + 
  geom_point(aes(x = dCT, y = No_Preds)) +
  geom_smooth(aes(x = dCT, y = No_Preds), method = "glm") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 12))

```
Looks like declines with increased distance from CT/increased Lat 


## Other spatial drivers?

### Can deal with longitude and distance from GBR at the same time
```{r}
# Longitude
predsum |> 
  ggplot() + 
  geom_point(aes(x = Site_long, y = No_Preds)) +
  geom_smooth(aes(x = Site_long, y = No_Preds), method = "glm") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 12))

# Distance from GBR
predsum |> 
  ggplot() + 
  geom_point(aes(x = dGBR, y = No_Preds), position = position_jitter(), shape = 21) +
  geom_smooth(aes(x = dGBR, y = No_Preds), method = "glm") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 12))

```

### Looks like some attenuation with distance offshore

# Other main spatial drivers around isolation and size:

## Reef_100
```{r}
predsum |> 
  ggplot() +
  #geom_point(aes(x = Reef_100, y = No_Preds)) +
  geom_smooth(aes(x = Reef_100, y = No_Preds), method = "glm") +
  theme_minimal() 
```


# Reef size
```{r}
predsum  |> 
  ggplot() +
  #geom_point(aes(x = Reef_1_Area, y = No_Preds)) +
  geom_smooth(aes(x = Reef_1_Area, y = No_Preds), method = "glm") +
  theme_minimal() +
  #scale_y_continuous(limits = c(0,12)) +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Reef size")
```

# Worth looking at some other habitat vars while we are here.

## Benthic complexity?
```{r}
predsum |> 
  #filter(Benthic_Complexity < 30) |> 
  mutate(Benthic_Complexity = factor(Benthic_Complexity, ordered = TRUE)) |> 
  group_by(Benthic_Complexity) %>% 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) |> 
  ggplot() + 
  geom_pointrange(aes(x = Benthic_Complexity, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Benthic_Complexity, y = mean)) +
  geom_line(aes(x = Benthic_Complexity, y = mean, group = 1)) +
  geom_smooth(aes(x = Benthic_Complexity, y = mean, group = 1), method = "glm") +
  theme_minimal() +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Benthic complexity score")
```

## Aspect?
```{r}
predsum |> 
  mutate(Aspect = factor(Aspect_scale, ordered = TRUE)) |> 
  group_by(Aspect) |> 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) |> 
  ggplot() + 
  geom_pointrange(aes(x = Aspect, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Aspect, y = mean)) +
  geom_line(aes(x = Aspect, y = mean, group = 1)) +
  geom_smooth(aes(x = Aspect, y = mean, group = 1), method = "glm") +
  theme_minimal() +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Aspect score (0-5)")
# OR

predsum  |> 
  group_by(Aspect_descriptive) |> 
  summarise_at(vars(No_Preds), list(mean = mean, SD = sd, SE = std.error)) |> 
  ggplot() + 
  geom_pointrange(aes(x = Aspect_descriptive, y = mean, ymin = mean-SE, ymax = mean+SE)) +
  geom_point(aes(x = Aspect_descriptive, y = mean)) +
  geom_line(aes(x = Aspect_descriptive, y = mean, group = 1)) +
  theme_minimal() +
  ylab("Abundance (no of predators ± SE)") +
  xlab("Aspect type")
```


# Temp - I think temp is probably co-linear with depth and tackled in a separate analysis.
```{r}
# predsum |> 
#   ggplot() + 
#   geom_point(aes(x = Temperature, y = No_Preds)) +
#   geom_smooth(aes(x = Temperature, y = No_Preds), method = "glm") +
#   theme_minimal() +
#   scale_y_continuous(limits = c(0, 8))
```


