---
title: "Plotting rarefaction curves"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```

# Prerequisites
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/helper_functions.R")
source("code/species_rich/sitedepth_rarefaction.R")

# OR
load("data/Rdata/predsum.Rdata")
load("data/Rdata/preds.Rdata")
load("data/Rdata/rmetadata.Rdata")
load("data/Rdata/species_rich_asymp.Rdata")
```



# Introduction

We currently have 33 sites over 13 reefs in our sampling design:
```{r}
# Total # sites = 33
rmetadata |> select(Site) |> distinct() |>  tally()
#rmetadata |> select(Reef_1) |> distinct()
```


And we have 3 depth bins, so if we sampled all of these we should have transects for 99 combinations.
However, we don't, as we didn't sample every Site_Depth combination:
```{r}
rmetadata |> select(Site_Depth) |> distinct()  |> tally()
```

We actually have 79 site~depth combinations (20 missing)



## In addition..

Sampling effort was not uniform among sites and depths, which adds a further level of imbalance to the story:

# Site/Reef ~ Depth transect combinations
```{r}
rmetadata |> group_by(Site, Depth) |> summarise(count = n()) |> pivot_wider(names_from = Site, values_from = count, values_fill = 0)
rmetadata |> group_by(Reef_1, Depth) |> summarise(count = n()) |> pivot_wider(names_from = Reef_1, values_from = count, values_fill = 0)
```

# We do have reasonable coverage across all "reef sizes", from small through to large. You'd expect lower numbers on the smallest reefs anyways, as you don't have as many places where you can sample:
```{r}
rmetadata |> 
  mutate(Reef_size = factor(case_when(Reef_1_Area < 3.5 ~ 'LQ',
                                      Reef_1_Area >= 3.5 & Reef_1_Area < 10.3 ~ 'Mean',
                                      Reef_1_Area >= 10.3 ~ 'UQ'))) |> 
  group_by(Reef_size) |> 
  summarise(count = n()) 
```

## And finally, within this unbalanced scenario, we have variance in the abundance of fishes observed, which will inevitably affect diversity estimates. 

All of this means we need a way to standardise our sampling and estimate actual levels of diversity at either reef or site level.

Rarefaction is a suitable approach that deals with unbalanced sampling. See this table for a summary of the predator counts by site/reef~depth. The count column is what is referred to in rarefaction as *"sampling depth"*
```{r}
sample_coverage <- 
  predsum |> 
  group_by(Reef_1, Site, Depth) |> 
  summarise(count = sum(No_Preds))  |> 
  arrange(-count)
sample_coverage
```

So while rarefaction is exactly designed to deal with these differences in sampling depth it won't cope with sites where there are 1 or less observations of fishes. That is 8 from our list of 79 site~depth combinations:
So we'll need to remove these from the analysis.


```{r}
sample_coverage |> 
  ggplot(aes(x = count)) +
  geom_histogram(binwidth = 2) +
  coord_cartesian(xlim = c(0, 75))
```



# Run rarefaction script
```{r}
source('code/diversity/depth_site_rarefaction.R')
```



## The files we need are buried in the raref object
```{r}
site_depth_rarefied$DataInfo
site_depth_rarefied$iNextEst$size_based
site_depth_rarefied$iNextEst$coverage_based 
site_depth_rarefied$AsyEst
```


# We just need the asymptotic estimates

```{r}
asym_diversity <- 
  env_vars |> 
  right_join(site_depth_rarefied$AsyEst |> 
  rename(Site_Depth = Assemblage)) |>
  mutate_if(is.character, as.factor)
```


# Size based curves and estimates
```{r}
# Species richness
rarefied_q0_sizebased <- 
  env_vars |> 
  right_join(site_depth_rarefied$iNextEst$size_based |> 
  rename(Site_Depth = Assemblage)) |>
  mutate_if(is.character, as.factor) |> 
  filter(Order.q == 0)

# Shannon
rarefied_q1_sizebased <- 
  env_vars |> 
  right_join(site_depth_rarefied$iNextEst$size_based |> 
  rename(Site_Depth = Assemblage)) |>
  mutate_if(is.character, as.factor) |> 
  filter(Order.q == 1)
```

# Extract asymptotes

# From the size-based rarefaction data
```{r}
asymp_rich1 <- 
  rarefied_q0_sizebased  |> 
  #ungroup |>  
  group_by(Site_Depth) |>  
  slice_max(qD) # Highest estimate of Species Richness

# From the asymptote tbl
asymp_rich2 <- asym_diversity %>% filter(Diversity == "Species richness")

asymp_check <- 
  asymp_rich2 %>%
  select(Site_Depth, Estimator) %>% 
  full_join(asymp_rich1 %>% 
              select(Site_Depth, qD, SC)) %>% 
  mutate(Diff = Estimator-qD) %>% 
  arrange(-Diff)

asymp_check
```
So where sample completeness is achieved, we have asymptotic richness estimates for all site~depths




```{r}
# From the coverage-based curve data
asymp_rich3 <- rarefied_q0_coveragebased %>%  group_by(Site_Depth) %>% slice_max(qD) # Highest estimate of Species Richness


asymp_check2 <- 
  asymp_rich2 %>%
  select(Site_Depth, Estimator) %>% 
  full_join(asymp_rich3 %>% 
              select(Site_Depth, qD, SC)) %>% 
  mutate(Diff = Estimator-qD) %>% 
  arrange(-Diff)
```



# Summary
We'll use the asym_rich2 dataframe (asymptotic diversity estimate object) as our raw species richness data file for the rest of the diversity analysis.
Note, that after the culling the data (via specifying minimum sampling depth), we lost quite a few sites (we are now down to 53 site~depth combinations instead of 71 that we started with) and we have lost Wreck Reef completely out of our dataset:

```{r}
asymp_rich2  |> select(Reef_1)  |>  distinct()
asymp_rich2  |> select(Site_Depth)  |>  distinct()
```
# Save file
```{r}
species_rich_asymp <- asymp_rich2 %>% 
  rename(SpecRich = Estimator)
#save(species_rich_asymp, file = "data/Rdata/species_rich_asymp.Rdata")
```





