---
title: "Predator assemblage composition using boral - region with data pooled at site-depth"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---



```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Prerequisites
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/helper_functions.R")
source("code/extras.R")
load("data/Rdata/predsum.Rdata")
load("data/Rdata/preds.Rdata")
#source("code/diversity/depth_site_diversity.R")
library(boral)
```



## Summarise species counts
```{r}
boral_speccount_TID <- 
  preds |> 
  group_by(Region, Reef_1, Site, Depth, Site_Depth, dCT, Binomial, T_ID) |> 
  dplyr::summarise(Count=sum(Number)) |> 
  ungroup()
```


## Convert into "wide" format diversity matrix - this is going to be our y variable
```{r}
boral_abund_TID <- 
  boral_speccount_TID |> 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup()
```


# And convert into matrix
```{r}
boral_matrix_TID <- 
  boral_abund_TID |> 
  select(!c(Region, Reef_1, Site, Depth, Site_Depth, dCT, T_ID)) |> 
  as.matrix()
```

## Store list of predictors - this is going to be our x variable
TID sampling scale

```{r}
env_boral_TID <- 
  preds |> 
  select(Region, Reef_1, Site,
         Reef_1_Area, Reef_100, Reef_size,
         Depth, Site_Depth,
         dCT, dGBR,
         T_ID) |> 
  distinct() |> 
  as.data.frame() |> 
  mutate(Region = fct_relevel(Region, "Central"))
```


# Set mcmc control parameters
```{r}
mcmc.control. <- list(n.burnin = 1000,
                      n.iteration = 7000,
                      n.thin = 30,
                      seed = 123)
```



# Fit boral model with region predictor variable

```{r}
#boral_TID_region_nb2 <- boral(boral_matrix_TID, X = env_boral_TID, family="negative.binomial",
#                   formula.X = ~ Region,
#                   #ranef.ids = env_boral_TID$Site,
#                   row.eff = "random",
#                   mcmc.control = mcmc.control.,
#                   lv.control = list(num.lv = 2),
#                   save.model = TRUE)
#save(boral_TID_region_nb2, file = "data/Rdata/boral_TID_region_nb2.rda")
load(file = "data/Rdata/boral_TID_region_nb2.rda")

```



# Model diagnostics

# Residuals
```{r}
plot(boral_TID_region_nb2)
```


# Check geweke diagnostics (chain convergence)
```{r}
# Raw geweke diagnostics
boral_TID_region_nb2$geweke.diag

#geweke.plot(boral_TID_region_nb2$jags.model)

#geweke.diag(get.mcmcsamples(boral_TID_region_nb2))

#get.mcmcsamples(boral_TID_region_nb2) %>% geweke.plot()

# Adjusted for multiple comparisons
#gew.pvals <- 2*pnorm(abs(unlist(boral_TID_region_nb2$geweke.diag[[1]])), lower.tail = FALSE) %>%  p.adjust(method = "holm")
#p.adjust(gew.pvals, method = "holm")
```


#

# For (slightly) better plotting we can use ggboral:
Compare to native plot
```{r}
gg_lvsplot(boral_TID_region_nb2, include = "both")
lvsplot(boral_TID_region_nb2, cex = 0.8)
```



# Can also "hack" ggboral to extract site and species scores:
First need to make the plots:
```{r}
sites_plot_TID_region <- gg_lvsplot(boral_TID_region_nb2, include = "objects", scaling = 1.2)
species_plot_TID_region <- gg_lvsplot(boral_TID_region_nb2, include = "attributes", scaling = 0.8)
```

# Compare to native plot
```{r}
species_plot_TID_region
sites_plot_TID_region
lvsplot(boral_TID_region_nb2, cex = 0.8, biplot = FALSE)
```

# Extract scores

## Species scores
```{r}
spec_scores_TID_region <- species_plot_TID_region$data
```

## Site scores
```{r}
site_scores_TID_region <- sites_plot_TID_region$data %>% 
  bind_cols(env_boral_TID)
```



# Plot
```{r}
ggplot(site_scores_TID_region, aes(x = lv1, y = lv2)) +
  geom_point(aes(colour = Region)) +
  ggforce::geom_mark_hull(aes(colour = Region),expand=0.02, concavity = 5) + 
  geom_point(data = spec_scores_TID_region) +
  theme_minimal() +
  facet_wrap(~ Region)
```




# To manually get at coefficients:

# So we'll make our own
```{r}
x <- boral_TID_region_nb2
```


```{r}
#alpha is an argument in lvsplot(); relates to scaling
alpha <- 0.5
```


```{r}
testcov <- x$lv.median %*% t(x$lv.coefs.median[, 2:3])

#singular value decom
do.svd <- svd(testcov, x$num.lv, x$num.lv)

svd(testcov)

#svd has components d, u and v

#some math related to rescaling
choose.lvs <- scale(do.svd$u *      matrix(do.svd$d[1:x$num.lv]^alpha, 
nrow = x$n, ncol = 2, byrow = T), center = T, scale = F)

#some more math related to rescaling
choose.lv.coefs <- scale(do.svd$v * matrix(do.svd$d[1:x$num.lv]^(1 - 
alpha), nrow = x$p, ncol = 2, byrow = T), center = T, 
                             scale = F)
```

```{r}
species_coeffs_TID_region <- 
colnames(boral_matrix_TID) %>% 
  as_tibble() %>% 
  rename(Species = value) %>% 
  bind_cols(choose.lv.coefs %>% 
              as_tibble())
```



# Check that our calculations produce the same scaled biplot:
```{r}
lvsplot(boral_TID_region_nb2, cex = 0.8, biplot = F, main = "real biplot")
lvsplot(boral_TID_region_nb2, cex = 0.8, biplot = T, main = "real biplot")
plot(choose.lvs, main = "hacked biplot")
```

# Plot properly

```{r}
#boral_ord_data <- 
  env_boral_TID |> 
  bind_cols(choose.lvs  |>  as.data.frame()) |> 
  ggplot(aes(x = V1, y = V2)) +
  geom_hline(yintercept = 0, lty = 11) +
  geom_vline(xintercept = 0, lty = 11) +
  geom_point(aes(colour = Region, shape = Region)) +
  geom_point(data = species_coeffs_TID_region) +
  ggforce::geom_mark_hull(aes(fill = Region), expand=0.02, concavity = 3) + 
  #ggforce::geom_mark_ellipse(aes(fill = Region), expand = 0, tol = 0.001) + 
  theme_bw() +
  facet_wrap(~ Region)
  
```





```{r}
#summary(boral_TID_region_nb2)
```




# Extract species coefficients
```{r}
# Mean values
Xcoefs_mean <-
  boral_TID_region_nb2$X.coefs.mean |> 
  as.matrix() |> 
  as.data.frame() |> 
  rownames_to_column(var = "Species")  |> 
  pivot_longer(cols = !Species, names_prefix = "Region", names_to = "Region", values_to = "Mean")

# Median values
Xcoefs_median <-
  boral_TID_region_nb2$X.coefs.median |> 
  as.matrix() |> 
  as.data.frame() |> 
  rownames_to_column(var = "Species")  |> 
  pivot_longer(cols = !Species, names_prefix = "Region", names_to = "Region", values_to = "Median")
    
# Lower CI
Xcoefs_lowerHPD <- 
boral_TID_region_nb2$hpdintervals$X.coefs |> 
  as.data.frame()  |> 
  rownames_to_column(var = "Species") |> 
  select(!5:7) |> 
  pivot_longer(cols = !Species, names_prefix = "Region", names_pattern = "(.*).lower", names_to = "Region", values_to = "lower.HPD")

# Upper CI
Xcoefs_upperHPD <- 
boral_TID_region_nb2$hpdintervals$X.coefs |> 
  as.data.frame()  |> 
  rownames_to_column(var = "Species") |> 
  select(!2:4) |> 
  pivot_longer(cols = !Species, names_prefix = "Region", names_pattern = "(.*).upper", names_to = "Region", values_to = "upper.HPD")
```

# Combine
```{r}
Xcoefs <- 
  Xcoefs_median |> 
  full_join(Xcoefs_lowerHPD) |> 
  full_join(Xcoefs_upperHPD) |> 
  mutate_if(is.character, as.factor)
```

# Now we could use some higher taxonomy info for plotting

# Create family order/index
```{r}
family_order <- 
  preds |> 
  select(Family) |> 
  distinct() |> 
  arrange(Family) |> 
  mutate(Fam_Ord = seq(1,22,1))
```

# Get family info from preds
```{r}
family_info <- 
  preds |> 
  select(Family, Binomial) |> 
  distinct() |> 
  rename(Species = Binomial) |> 
  left_join(family_order) # And add the order back in 
```


# Extract just the significant differences
```{r}
sigXcoefs <- Xcoefs |> 
  left_join(family_info) |> 
      left_join(family_info) |> 
  mutate(TestA = ifelse(lower.HPD < 0, -1, 0),
         TestB = ifelse(upper.HPD > 0, 1, 0),
         Flag = TestA+TestB) |> 
  filter(Flag != 0) |> 
  mutate(Species = fct_reorder(Species, Fam_Ord)) 
```




# Plot significant taxa
```{r}
sigXcoefs |> 
  ggplot(aes(x = Species, y = Median)) +
  geom_pointrange(aes(ymin = lower.HPD, ymax = upper.HPD, colour = Family)) +
  geom_point(aes(fill = Family), pch = 21, colour = "black", size = 3) +
  geom_hline(yintercept = 0, lty = 11) +
  scale_fill_viridis_d() +
  scale_colour_viridis_d() +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x = element_blank()) +
  facet_wrap(~ Region)
```

# What we have just made is a pretty version of this, which is downright horrible:
```{r}
coefplot <- sapply(colnames(boral_nb2$X), coefsplot, object = boral_nb2)
```

# If you want to dig into spp correlations (we don't, right now):
```{r}
get.enviro.cor(boral_nb2)
```



# To plot significant taxa from mvabund

```{r}
sigXcoefs2 <- 
  Xcoefs |> 
  left_join(regionTID_uni_contrasts |> 
              rename(Species = Binomial)) %>% 
  filter(P <= 0.05)

```










# See if there are differences between reef sizes

```{r}
#boral_nb3 <- boral(depth_site_matrix, X = env_boral, family="negative.binomial",
#                   formula.X = ~ Reef_size,
#                   #ranef.ids = env_boral$Site,
#                   row.eff = "random",
#                   mcmc.control = mcmc.control.,
#                   lv.control = list(num.lv = 2),
#                   save.model = TRUE)

#save(boral_nb3, file = "data/Rdata/boral_nb3.rda")
load(file = "data/Rdata/boral_nb3.rda")

```


# Check geweke diagnostics (chain convergence)
```{r}
# Raw geweke diagnostics
#boral_nb3$geweke.diag
#geweke.plot(boral_nb3$jags.model)
#geweke.diag(get.mcmcsamples(boral_nb2))
#get.mcmcsamples(boral_nb2) %>% geweke.plot()
# Adjusted for multiple comparisons
#gew.pvals <- 2*pnorm(abs(unlist(boral_nb2$geweke.diag[[1]])), lower.tail = FALSE) %>%  p.adjust(method = "holm")
#p.adjust(gew.pvals, method = "holm")
```


# Again we'll extract coefficients and make our own plot
```{r}
x <- boral_nb3
```


```{r}
#alpha is an argument in lvsplot(); relates to scaling
alpha <- 0.5
```


```{r}
testcov <- x$lv.median %*% t(x$lv.coefs.median[, 2:3])

#singular value decom
do.svd <- svd(testcov, x$num.lv, x$num.lv)

#svd has components d, u and v

#some math related to rescaling
choose.lvs <- scale(do.svd$u *      matrix(do.svd$d[1:x$num.lv]^alpha, 
                                           nrow = x$n, ncol = 2, byrow = T), center = T, scale = F)

#some more math related to rescaling
choose.lv.coefs <- scale(do.svd$v * matrix(do.svd$d[1:x$num.lv]^(1 - 
                                                                   alpha), nrow = x$p, ncol = 2, byrow = T), center = T, 
                         scale = F)
```


# Check that our calculations produce the same scaled biplot:
```{r}
lvsplot(boral_nb3, cex = 0.8, biplot = F, main = "real biplot")
plot(choose.lvs, main = "hacked biplot")
```

Looks good, can plot properly now


```{r}
#boral_ord_data <- 
env_boral_SD |> 
  bind_cols(choose.lvs  |>  as.data.frame()) |> 
  ggplot(aes(x = V1, y = V2)) +
  geom_hline(yintercept = 0, lty = 11) +
  geom_vline(xintercept = 0, lty = 11) +
  geom_point(aes(colour = Reef_size, shape = Reef_size)) +
  ggforce::geom_mark_hull(aes(fill = Reef_size), expand=0.02, concavity = 3) + 
  #ggforce::geom_mark_ellipse(aes(fill = Region), expand = 0, tol = 0.001) + 
  theme_bw() +
  facet_wrap(~ Reef_size)

```





```{r}
summary(boral_nb3)
```




# Extract species coefficients
```{r}
# Mean values
Xcoefs_mean <-
  boral_nb3$X.coefs.mean |> 
  as.matrix() |> 
  as.data.frame() |> 
  rownames_to_column(var = "Species")  |> 
  pivot_longer(cols = !Species, names_prefix = "Reef_size", names_to = "Reef_size", values_to = "Mean")

# Median values
Xcoefs_median <-
  boral_nb3$X.coefs.median |> 
  as.matrix() |> 
  as.data.frame() |> 
  rownames_to_column(var = "Species")  |> 
  pivot_longer(cols = !Species, names_prefix = "Reef_size", names_to = "Reef_size", values_to = "Median")

# Lower CI
Xcoefs_lowerHPD <- 
  boral_nb3$hpdintervals$X.coefs |> 
  as.data.frame()  |> 
  rownames_to_column(var = "Species") |> 
  select(!4:5) |> 
  pivot_longer(cols = !Species, names_prefix = "Reef_size", names_pattern = "(.*).lower", names_to = "Reef_size", values_to = "lower.HPD")

# Upper CI
Xcoefs_upperHPD <- 
  boral_nb3$hpdintervals$X.coefs |> 
  as.data.frame()  |> 
  rownames_to_column(var = "Species") |> 
  select(!2:3) |> 
  pivot_longer(cols = !Species, names_prefix = "Reef_size", names_pattern = "(.*).upper", names_to = "Reef_size", values_to = "upper.HPD")
```

# Combine
```{r}
Xcoefs <- 
  Xcoefs_median |> 
  full_join(Xcoefs_lowerHPD) |> 
  full_join(Xcoefs_upperHPD) |> 
  mutate_if(is.character, as.factor)
```

# Now we could use some higher taxonomy info for plotting

# Create family order/index
```{r}
family_order <- 
  preds |> 
  select(Family) |> 
  distinct() |> 
  arrange(Family) |> 
  mutate(Fam_Ord = seq(1,22,1))
```

# Get family info from preds
```{r}
family_info <- 
  preds |> 
  select(Family, Binomial) |> 
  distinct() |> 
  rename(Species = Binomial) |> 
  filter(!(str_detect(Species, "Aphar") & str_detect(Family, "Carang")))  |> # This also removes the 'NA' taxon
  mutate(Family = replace(Family, str_detect(Species, "Epinephelus "), "Serranidae")) |> 
  mutate(Family = replace(Family, str_detect(Species, "Bodianus "), "Labridae"))  |> 
  left_join(family_order) |>   # And add the order back in 
  droplevels()

family_info  %$% summary(factor(Species))

```


# Extract just the significant differences
```{r}
sigXcoefs <- Xcoefs |> 
  left_join(family_info) |> 
  mutate(TestA = ifelse(lower.HPD < 0, -1, 0),
         TestB = ifelse(upper.HPD > 0, 1, 0),
         Flag = TestA+TestB) |> 
  filter(Flag != 0) |> 
  mutate(Species = fct_reorder(Species, Fam_Ord)) 

sigXcoefs  %$% summary(factor(Species))
```




# Plot significant taxa
```{r fig.width = 160/25.4, fig.height=100/25.4}
taxa_diffs <- 
sigXcoefs |> 
  mutate(Reef_size1 = case_when(grepl("Mean", Reef_size) ~ "Mid-sized reefs",
                                grepl("UQ", Reef_size) ~ "Large reefs")) |> 
  ggplot(aes(x = Species, y = Median)) +
  geom_pointrange(aes(ymin = lower.HPD, ymax = upper.HPD), colour = "black") +
  geom_point(aes(fill = Family), pch = 21, colour = "black", size = 3) +
  geom_hline(yintercept = 0, lty = 11) +
  scale_fill_viridis_d(guide = guide_legend(reverse = TRUE)) +
 # scale_colour_viridis_d() +
  scale_y_continuous("Latent variable coef. contrasts (median ± 95% CI)") +
  coord_flip() +
  theme_bjc() +
  theme(axis.text.x = element_text()) +
  facet_wrap(~ ~fct_rev(Reef_size1)) +
  theme(axis.text.y = element_text(face = 'italic'))

taxa_diffs
```

## Next, we'll marry this up with lob_preds to demonstrate abundance with depth for each taxon

# Extract significant taxa
```{r}
sig_taxa <- 
sigXcoefs |> 
  select(Species) %>% 
  distinct()  |> 
  mutate(Species = as.character(Species)) |> 
  arrange(Species)
sig_taxa
```


```{r}
sig_preds <- 
  sig_taxa |> 
  select(Species) %>% 
  distinct() %>% 
  rename(Binomial = Species) %>% 
  left_join(preds)
```


```{r}
sig_predsum <- 
   sig_preds  %>% 
   group_by(Reef_size, Reef_1, Binomial) %>%
   summarise(No_Preds =sum(Number)) %>%  # All good - 514
   ungroup() %>% 
  complete(Reef_size, Binomial, Reef_1) %>% 
  mutate(No_Preds = replace_na(No_Preds, 0))
```

```{r}
sig_predsum2 <- sig_predsum %>% 
  group_by(Reef_size, Binomial) %>%
  summarise(mean = mean(No_Preds),
            sd = sd(No_Preds, na.rm = TRUE),
            SE = std.error(No_Preds, na.rm = TRUE)) %>% 
  arrange(-mean) %>% 
  arrange(Reef_size) 
```




```{r fig.width=160/25.4, fig.height=80/25.4}
sig_abun_plot <- 
sig_predsum2 %>% 
  ggplot(aes(x = Reef_size, y = mean, fill = Reef_size)) +
  geom_line(aes(group = 1)) +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE)) +
  geom_point(pch = 21, size = 2) +
  scale_y_continuous(name = "Mean abundance ± SE") +
  theme_bjc() +
  theme(legend.title = element_blank(),
        legend.position = c(0.5, 0.07),
        legend.direction = "horizontal",
        axis.text = element_blank(),
        strip.text.x = element_text(face = "italic")) +
  facet_wrap(~ Binomial, nrow = 3, scales = "free_y")

sig_abun_plot
```








```{r}
coefplot <- sapply(colnames(boral_nb3$X), coefsplot, object = boral_nb3)

sapply(colnames(boral_nb3$X), coefsplot, object = boral_nb3)
```



```{r}
ggsave(filename = "output/taxa_diffs_reefsize.pdf", plot = taxa_diffs, height = 100, width = 160, device = cairo_pdf, units = "mm", dpi = 300)
```











