---
title: "boral modelling predator assemblage: LV only"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---



```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Prerequisites
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/extras.R")
load("data/Rdata/preds.Rdata")
library(boral)
library(ggboral)
```


## Summarise species counts
```{r}
boral_speccount_SD <- 
  preds |> 
  group_by(Region, Reef_1, Site, Depth, Site_Depth, dCT, Binomial) |> 
  dplyr::summarise(Count=sum(Number)) |> 
  ungroup()
```


## Convert into "wide" format diversity matrix - this is going to be our y variable
```{r}
boral_abund_SD <- 
  boral_speccount_SD |> 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup()
```


# And convert into matrix
```{r}
boral_matrix_SD <- 
  boral_abund_SD |> 
  select(!c(Region, Reef_1, Site, Depth, Site_Depth, dCT)) |> 
  as.matrix()
```

## Store list of predictors - this is going to be our x variable
At depth_site pooling
```{r}
env_boral_SD <- 
  preds |> 
  select(Region, Reef_1, Site,
         Reef_1_Area, Reef_100, Reef_size,
         Depth, Site_Depth,
         dCT, dGBR
         ) |> 
  distinct() |> 
  as.data.frame() |> 
  mutate(Region = fct_relevel(Region, "Central"))
```


# Set mcmc control parameters
```{r}
mcmc.control. <- list(n.burnin = 1000,
                      n.iteration = 7000,
                      n.thin = 30,
                      seed = 123)
```


# Fit model
We can construct a latent variable only model, with no env predictors involved, which will allow us to construct an ordination:
```{r}
#boral_sd_nb1 <- boral(boral_matrix_SD, family="negative.binomial",
#                   mcmc.control = mcmc.control.,
#                   lv.control = list(num.lv = 2))

#save(boral_sd_nb1, file = "data/Rdata/boral_sd_nb1.rda")
load(file = "data/Rdata/boral_sd_nb1.rda")
```


# Model diagnostics

# Residuals
```{r}
plot(boral_sd_nb1)
```

# Chain convergence 
Via geweke diagnostics
```{r}
# Raw geweke diagnostics
boral_sd_nb1$geweke.diag
geweke.plot(boral_sd_nb1$jags.model)
geweke.diag(get.mcmcsamples(boral_sd_nb1))
get.mcmcsamples(boral_sd_nb1) %>% geweke.plot()

# Adjusted for multiple comparisons
gew.pvals <- 2*pnorm(abs(unlist(boral_sd_nb1$geweke.diag[[1]])), lower.tail = FALSE)
p.adjust(gew.pvals, method = "holm")
```


# Model investigation
Doesn't really have anything in here as it is just the latent variables behind this model fit (unconstrained)
```{r}
#tidyboral(boral_sd_nb1)
#attributes(boral_sd_nb1)
```


# From this model we can make a basic (horrible) biplot:
```{r}
lvsplot(boral_sd_nb1, cex = 0.8)
```

# Or with species names removed
```{r}
lvsplot(boral_sd_nb1, cex = 0.8, biplot = FALSE)
```

# For (slightly) better plotting we can use ggboral:
Compare to native plot
```{r}
gg_lvsplot(boral_sd_nb1, include = "both")
lvsplot(boral_sd_nb1, cex = 0.8)
```
Matches up well

# Can also "hack" ggboral to extract site and species scores:
First need to make the plots:
```{r}
sites_plot <- gg_lvsplot(boral_sd_nb1, include = "objects")
species_plot <- gg_lvsplot(boral_sd_nb1, include = "attributes")
```

# Compare to native plot
```{r}
species_plot
sites_plot
lvsplot(boral_sd_nb1, cex = 0.8, biplot = FALSE)
```

# Extract scores
Species scores
```{r}
spec_scores <- species_plot$data
```




```{r}
site_scores <- sites_plot$data %>% 
  bind_cols(env_boral_SD)
```



#

```{r}
ggplot(site_scores, aes(x = lv1, y = lv2)) +
  geom_point(aes(colour = Region)) +
  ggforce::geom_mark_hull(aes(colour = Region),expand=0.02, concavity = 5) + 
  geom_point(data = spec_scores) +
  theme_minimal() +
  facet_wrap(~ Region)
```




















## More info ##
https://rpubs.com/brouwern/boralnotes