---
title: "Predator assemblage composition - differences among reef sizes - TID scale"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Prerequisites
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
#source("code/helper_functions.R")
source("code/extras.R")
#load("data/Rdata/predsum.Rdata")
load("data/Rdata/preds.Rdata")
#load("data/Rdata/rmetadata.Rdata")
#source("code/diversity/depth_site_diversity.R")
```




## Store list of predictors we'll be interested in at the end.
```{r}
env_vars_size_TID <- 
  preds |> 
  #filter(Region != "South") |> 
  select(Reef_size, T_ID) |> 
  distinct()
```



## Summarise species accounts by depth bin and site:
```{r}
sizeTID_speccount_mva <- 
  preds |> 
#  filter(Region != "South") |> 
  group_by(T_ID, Binomial) |> # Just the grouping variables that we are interested in here
  dplyr::summarise(Count=sum(Number)) |> 
  ungroup()
```




## Convert into "wide" format diversity matrix
```{r}
sizeTID_matrix_mva <- 
  sizeTID_speccount_mva |> 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup() 
```


## Turn into mvabund object
```{r}
predabund <- mvabund(sizeTID_matrix_mva[-1]) 
```

## Check out mean-variance relationship - should be 1-1 as we are working with count data
```{r}
par(predabund)
meanvar <- meanvar.plot(predabund) 
```

# EDA - can do a basic overview of abundance ~ predictor variable
Won't work in Rmd!! Need to copy to script or console
```{r}
#plot(predabund ~ env_vars_size_TID$Reef_size)
```


#  Model fitting
## Notes:
- mvabund cannot handle random factors (yet) so a resampling approach used instead (see below)
- There is a pattern in the Poisson model residuals (not surprising, given the ZI in these data) so have commented out

```{r}
# Neg binom model options
reefsizeTID_mod <- manyglm(predabund ~ env_vars_size_TID$Reef_size, family="negative_binomial")

```



# Model validation 
# In mvabund the only way to check this is plotting residuals vs fitted values
```{r}
plot.manyglm(reefsizeTID_mod, which = 1) # 
plot.manyglm(reefsizeTID_mod, which = 2) # 
plot.manyglm(reefsizeTID_mod, which = 3) # L
```



# Summaries
```{r}
# Summary of model
#reefsizeTID_modsum <- summary(reefsizeTID_mod)

# Save as these take a while to run
#save(reefsizeTID_modsum, file = "data/Rdata/reefsizeTID_modsum.Rdata") # 

# Load when required
load(file = "data/Rdata/reefsize_modsum.Rdata") # 

# Inspect
reefsize_modsum
```


# Pairwise comparisons
```{r}
#reefsizeTID_aov <- anova.manyglm(reefsizeTID_mod, test = 'LR', pairwise.comp = env_vars_size_TID$Reef_size,  p.uni="adjusted")
```

# These anovas are very slow to run -save results for use later
```{r}
#save(reefsizeTID_aov, file = "data/Rdata/reefsizeTID_aov.Rda") # 
```

# Reload if/when required
```{r}
load(file = "data/Rdata/reefsizeTID_aov.Rda")
```

# Results
```{r}
summary(reefsizeTID_aov)
```


# Contrasts
```{r}
reefsize_multi_contrasts <- 
  reefsizeTID_aov$pairwise.comp.table  |> 
              rownames() |> 
              as_tibble() |> 
              mutate(Contrast = value, .keep = "unused") |> 
  bind_cols(reefsizeTID_aov$pairwise.comp.table  |> 
  as_tibble())

reefsize_multi_contrasts
```


```{r}

reefsize_uni_contrasts <- 
  t(reefsizeTID_aov$uni.test) %>% 
  as.data.frame() %>% 
  rownames_to_column("Binomial") %>% 
  rename(LRT = `env_vars_size_TID$Reef_size`) %>% 
  select(!`(Intercept)`) %>% 
  left_join(  
  t(reefsizeTID_aov$uni.p) %>% 
  as.data.frame() %>% 
  rownames_to_column("Binomial") %>% 
  rename(P = `env_vars_size_TID$Reef_size`) %>% 
  select(!`(Intercept)`)) %>% 
  mutate(Binomial = str_replace(Binomial, "\\.", " ")) %>% 
  arrange(-LRT)

reefsize_uni_contrasts

```


```{r}
reefsize_sig_preds <- 
  reefsize_uni_contrasts |> 
  filter(P <= 0.05) %>% 
  left_join(preds)
```


```{r}
reefsize_sig_predsum <- 
   reefsize_sig_preds  %>% 
   group_by(Reef_size, Reef_1, Binomial) %>%
   summarise(No_Preds = sum(Number)) %>%  # All good - 514
   ungroup() %>% 
  complete(Reef_size, Reef_1, Binomial) %>% 
  mutate(No_Preds = replace_na(No_Preds, 0)) %>% 
  group_by(Reef_size, Binomial) %>%
  summarise(mean = mean(No_Preds),
            sd = sd(No_Preds, na.rm = TRUE),
            SE = std.error(No_Preds, na.rm = TRUE)) %>% 
  arrange(-mean) %>% 
  arrange(Reef_size) 
```



```{r fig.width=80/25.4, fig.height=70/25.4}
reefsize_sig_predsum_plot <- 
reefsize_sig_predsum %>% 
  ggplot(aes(x = Reef_size, y = mean, fill = Reef_size)) +
  geom_line(aes(group = 1)) +
  geom_linerange(aes(ymin = mean - SE, ymax = mean + SE)) +
  geom_point(mapping = aes(shape = Reef_size), size = 3) +
  scale_y_continuous(name = expression("Mean abund. (.150m"^{2}~" ± SE)"), expand = c(0.1, 0.1)) +
  scale_x_discrete(labels =  c("Small", "Mid-size", "Large"), name = "Reef size") +
  scale_fill_manual(labels =  c("Small", "Mid-size", "Large"), values = c("lightskyblue", "dodgerblue3", "darkblue")) +
  scale_shape_manual(labels =  c("Small", "Mid-size", "Large"), values = c(24, 22, 25)) +
  #scale_size_manual(labels =  c("Small", "Mid-size", "Large"), values = c(3, 5, 10)) +
  theme_bjc() + 
  theme(legend.position = "none",
        legend.title = element_blank(),
        strip.text = element_text(size = 10, face = "italic"),
        axis.title.x = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = "black")) +
  facet_wrap(~ Binomial, nrow = 2, scales = "free_y")

reefsize_sig_predsum_plot

```

