---
title: "boral modelling predator assemblage: LV only - transect as sampling unit"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---



```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Prerequisites
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/packages.R")
source("code/extras.R")
load("data/Rdata/preds.Rdata")
library(boral)
library(ggboral)
```


## Summarise species counts
```{r}
boral_speccount_TID <- 
  preds |> 
  group_by(Region, Reef_1, Site, Depth, Site_Depth, dCT, Binomial, T_ID) |> 
  dplyr::summarise(Count=sum(Number)) |> 
  ungroup()
```


## Convert into "wide" format diversity matrix - this is going to be our y variable
```{r}
boral_abund_TID <- 
  boral_speccount_TID |> 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0) |> 
  dplyr::select(!`NA`) |>  # Get rid of column where the sp name ('Binomial') had been NA
  ungroup()
```


# And convert into matrix
```{r}
boral_matrix_TID <- 
  boral_abund_TID |> 
  select(!c(Region, Reef_1, Site, Depth, Site_Depth, dCT, T_ID)) |> 
  as.matrix()
```

## Store list of predictors - this is going to be our x variable
TID sampling scale

```{r}
env_boral_TID <- 
  preds |> 
  select(Region, Reef_1, Site,
         Reef_1_Area, Reef_100, Reef_size,
         Depth, Site_Depth,
         dCT, dGBR,
         T_ID) |> 
  distinct() |> 
  as.data.frame() |> 
  mutate(Region = fct_relevel(Region, "Central"))
```


# Set mcmc control parameters
```{r}
mcmc.control. <- list(n.burnin = 1000,
                      n.iteration = 7000,
                      n.thin = 30,
                      seed = 123)
```


# Fit model
We can construct a latent variable only model, with no env predictors involved, which will allow us to construct an ordination:
```{r}
#boral_TID_LV <- boral(boral_matrix_TID, family="negative.binomial",
#                   mcmc.control = mcmc.control.,
#                   lv.control = list(num.lv = 2),
#                   save.model = TRUE)

#save(boral_TID_LV, file = "data/Rdata/boral_TID_LV.rda")
load(file = "data/Rdata/boral_TID_LV.rda")
```


# Model diagnostics

# Residuals
```{r}
plot(boral_TID_LV)
```

# Chain convergence 
Via geweke diagnostics
```{r}
# Raw geweke diagnostics
boral_TID_LV$geweke.diag
#geweke.plot(boral_TID_LV$jags.model)
#geweke.diag(get.mcmcsamples(boral_TID_LV))
#get.mcmcsamples(boral_TID_LV) %>% geweke.plot()
#
# Adjusted for multiple comparisons
#gew.pvals <- 2*pnorm(abs(unlist(boral_TID_LV$geweke.diag[[1]])), lower.tail = FALSE)
#p.adjust(gew.pvals, method = "holm")
```


# Model investigation
Doesn't really have anything in here as it is just the latent variables behind this model fit (unconstrained)
```{r}
#tidyboral(boral_TID_LV)
#attributes(boral_TID_LV)
```


# From this model we can make a basic (horrible) biplot:
```{r}
lvsplot(boral_TID_LV, cex = 0.8)
```

# Or with species names removed
```{r}
lvsplot(boral_TID_LV, cex = 0.8, biplot = FALSE)
```

# For (slightly) better plotting we can use ggboral:
Compare to native plot
```{r}
gg_lvsplot(boral_TID_LV, include = "both")
lvsplot(boral_TID_LV, cex = 0.8)
```
Matches up well

# Can also "hack" ggboral to extract site and species scores:
First need to make the plots:
```{r}
sites_plot_LV <- gg_lvsplot(boral_TID_LV, include = "objects")
species_plot_LV <- gg_lvsplot(boral_TID_LV, include = "attributes")
```

# Compare to native plot
```{r}
species_plot_LV
sites_plot_LV
lvsplot(boral_TID_LV, cex = 0.8, biplot = FALSE)
```

# Extract scores
Species scores
```{r}
spec_scores_LIV <- species_plot_LV$data

sig_spec_scores_LIV <- species_plot_LV$data |>  
  rename(Binomial = label)  |> 
  left_join(reefsize_uni_contrasts) |>  # Need to run the mvabund Rmd first
  filter(P <= 0.05)
```

Site scores
```{r}
site_scores_LV <- sites_plot_LV$data %>% 
  bind_cols(env_boral_TID)
```




# Ordination by region
```{r fig.width=180/25.4, fig.height=170/25.4}
ggplot(site_scores_LV, aes(x = lv1, y = lv2)) +
  geom_point(data = sig_spec_scores_LIV) +
  ggforce::geom_mark_hull(aes(colour = Region), expand=0.02, concavity = 5) + 
  geom_point(aes(colour = Region)) +
  theme_minimal()
  facet_wrap(~ Region)
```


# Ordination by reef size

# Filtering out non-interesing taxa (those near centre of ordination)
Will do with 2 x SE (0.68*2)
```{r}
circle.prob <- 0.68 # 1 x Std Error
r <- 3*sqrt(qchisq(circle.prob, df = 2)) * prod(colMeans(site_scores_LV[,1:2]^2))^(1/4) # Using sin and cosin to make a circle
theta <- c(seq(-pi, pi, length = 50), seq(pi, -pi, length = 50))
circle <- data.frame(PC1 = r * cos(theta), PC2 = r * sin(theta))
```

```{r}
sp_posn <- read_excel("data/sp_posn.xlsx")
```


```{r}
sig_spec_scores_LIV2 <- species_plot_LV$data |>  
  rename(Species = label)  |> 
  mutate(Flag = factor(ifelse(sqrt(lv1^2 + lv2^2) > r, 1, 0)))  |> 
  filter(Flag != 0 | Species == "Monotaxis heterodon") |> 
  left_join(family_info) |> 
  arrange(Family)  |> 
  left_join(read_excel("data/sp_posn.xlsx")) |> 
  mutate(Number = seq(1,12,1))

# How many fams in here?
sig_spec_scores_LIV2 |> 
  select(Family) |> 
  distinct()
```


```{r fig.width=180/25.4, fig.height=110/25.4}
boral_ord_size <- 
ggplot(site_scores_LV, aes(x = lv1, y = lv2)) +
  geom_segment(data=sig_spec_scores_LIV2, aes(y=0, x=0, yend=lv2, xend=lv1), arrow=arrow(length=unit(0.3,'lines')), color='grey30') +
  stat_ellipse(aes(fill = Reef_size, colour = Reef_size, group = Reef_size), geom = "polygon", type = "norm" , level = 0.95, alpha = 0.6) +
  geom_point(data=sig_spec_scores_LIV2, aes(shape = Family), fill = "grey40", size = 3) +
  geom_point(aes(fill = Reef_size), pch = 21, size = 2, alpha = 0.5) +
  #geom_path(data = circle, aes(y=PC2,x=PC1), color = 'black', fill = "white", linewidth = 1/2, alpha = 1/3) +
  geom_text(data=sig_spec_scores_LIV2, aes(x = spx, y = spy, label = Species), fontface = "italic", show.legend = FALSE) +
  #stat_ellipse(aes(colour = Reef_size, group = Reef_size), fill = "white", geom = "polygon", type = "norm" , level = 0.95, alpha = 1) +
  geom_hline(yintercept=0, linetype='dotted') +
  geom_vline(xintercept=0, linetype='dotted') +  
  scale_fill_viridis_d() +
  scale_colour_viridis_d() +
  scale_shape_manual(values = c(21,22,24,25)) +
  scale_y_continuous("L.V.2") +
  scale_x_continuous("L.V.1", limits = c(-4.5, 5.1)) +
  theme_bjc() +
  theme(panel.background = element_rect(fill = "transparent", colour = "black"),
        legend.position = "right") 
  
boral_ord_size
```

# Save

```{r}
ggsave(plot = boral_ord_size, filename = "output/boral_ord_size.jpg", height = 110, width = 170, device = "jpeg", units = "mm", dpi = 300)
```

















## More info ##
https://rpubs.com/brouwern/boralnotes