---
title: " Mapping CSMP Overview"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    #fig_height: 
    fig_width: 140/25.4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---
## Created: 3rd Oct 2023 BJC ##



```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```



# Load required packages
```{r packages, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
source('code/packages.R')
library(sf)
library(rio)
library(ggsn)

```







# First need a base map of the coastal area

## Load Australia map object from ggplot's world map
Note: could use whole world map but the file will be 50 x smaller if we filter now (0.1 vs ~5MB), and thus quicker to plot and work with
```{r world-map-load}
aus <- map_data("world") %>% 
  filter(region == 'Australia')
```

# Load reef labels
```{r}
reef_labels <- 
  read_csv("data/reef_locs.csv")
```


# Add globe and scale bar
```{r}
sourceRmd::sourceRmd("code/make_globe.Rmd")
```

## Load in shape files 
Need to have the .shp and the .shx file in the same folder
# CS
```{r load-shp-sf, message=FALSE, warning=FALSE}
cs_dryreef <- read_sf("data/3dgbr_geomorph/shape/coralsea_dryreef.shp")
cs_reef <- read_sf("data/3dgbr_geomorph/shape/coralsea_reef.shp")
cs_cays <- read_sf("data/3dgbr_geomorph/shape/coralsea_cay.shp")
```

# GBR
```{r}
gbr_feats <- st_read("data/3dgbr_geomorph/shape/gbr_features.shp")
gbr_shelf <- st_read("data/3dgbr_geomorph/shape/gbr_shelf.shp") # Don't need to filter as the file is already as small as it can be
```
## Basic map trimmed down to just GBR and Coral Sea area
- Could specify detailed aes here but will save that step for the final maps to keep this early code easy to follow
```{r}
raw_map <- 
  aus %>% 
  ggplot() +
  geom_sf(data = gbr_feats, fill = 'grey60', colour = "transparent", size = .001) +
  geom_polygon(aes(x = long, y = lat, group = group), fill = "grey40", colour = 'black', linewidth = 0.3) +
  coord_sf(ylim = c(-23, -9.5), xlim = c(143, 170)) 
print(raw_map)
```

## Whole CSMP general area
```{r}
#overview_map <- 
raw_map +
  #geom_sf(data = cs_reef, fill = 'gray85', colour = 'black', size = 0.1) + # Can use fill = on polygon but not path
  geom_sf(data = cs_dryreef, fill = 'gray60', size = 0.1) +
  geom_sf(data = cs_cays, fill = 'black', colour = 'black', size = 0.1) +
  geom_point(data = reef_locs, mapping = aes(x = Site_long, y = Site_lat)) +
  coord_sf(ylim = c(-23, -9.5), xlim = c(143, 157)) +
       theme(axis.title = element_blank(), 
        #axis.text = element_blank(),
        #axis.ticks = element_blank(),
        #panel.grid = element_line(colour = "#ECEFF1"),
        panel.border = element_rect(colour = "black", fill = "transparent"),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", colour = "transparent"))
```



# Add detail to base map
```{r}
base_map <- 
  raw_map +
  geom_sf(data = cs_dryreef, fill = 'gray60', size = 0.1) +
  geom_sf(data = cs_cays, fill = 'black', colour = 'black', size = 0.1) +
  geom_segment(data = reef_labels, mapping = aes(x = Site_long, xend = Label_long, y = Site_lat, yend = Label_lat)) +
  geom_point(data = reef_labels, mapping = aes(x = Label_long, y = Label_lat, group = Region, fill = Region, shape = Region), size = 6) +
  geom_text(data = reef_labels, mapping = aes(x = Label_long, y = Label_lat, label = Abbr, group = Region, shape = Region), family = "Arial", size = 3) +
  scale_shape_manual(values = c(21,22,23,24)) +
  coord_sf(ylim = c(-23, -9.5), xlim = c(143, 157)) +
       theme(axis.title = element_blank(),
             #axis.text = element_blank(),
             #axis.ticks = element_blank(),
             #panel.grid = element_line(colour = "#ECEFF1"),
             panel.border = element_rect(colour = "black", fill = "transparent"),
             panel.background = element_rect(fill = "transparent"),
             plot.background = element_rect(fill = "transparent", colour = "transparent"))

base_map
```








```{r fig.width= 160/25.4, fig.height=160/25.4}
reef_map <- base_map +
  annotation_custom(ggplotGrob(globe), xmin = 143, xmax = 147, ymin = -23, ymax = -19) +
  scalebar(x.min = 150, x.max = 154, y.min = -10, y.max = -9, dist = 250, dist_unit = "km", transform = TRUE, model = "WGS84", # Geo properties - the xmax and ymax don't do anything
          # location = 'bottomright', anchor = c(x = 149, y = -18.1),                                # Position of bar
           family = "Arial", st.size = 3, st.dist = 0.3,                                            # Text properties
           border.size = 0.3, height = 0.2, st.bottom = TRUE) +                                     # Bar properties
  cowplot::draw_image('data/North_Pointer.jpg', x = 156, y = -11, scale = 1.6, hjust = 0.5) +
#  annotation_custom(ggplotGrob(globe), xmin = 146, xmax = 148, ymin = -16, ymax = -14.5) 
  geom_rect(aes( xmin = 144.9, xmax = 145.3, ymin = -21.45, ymax = -21), linewidth = 0.01, color = "red", fill = NA) +
    guides(fill = "none", shape = "none") +
  theme_bjc()
reef_map
```






```{r}
ggsave(filename = "output/reef_map.pdf", plot = reef_map, height = 160, width = 160, device = cairo_pdf, units = "mm", dpi = 300)
```




















