---
title: "Diversity modelling - Isolation"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```

```{r}
source('code/packages.R')
source('code/helper_functions.R')
```


# We'll start with the aysymptotic species richness estimates from iNEXT.
```{r}
load(file = "data/Rdata/species_rich_asymp.Rdata")
```

```{r}
glimpse(species_rich_asymp)
```

# Check distribution of response variable
```{r}
species_rich_asymp |> 
  ggplot() +
  geom_boxplot(aes(y = log(SpecRich)))
```

Definitely need a model with a log link to handle the response distribution.


As with the abundance modelling, the first thing we are interested in is the combined effect of isolation on diversity and we already have these variables in the data:

-------------     ------------------------------------------------------------------------------
**Reef_1_Area**       Size of the reef in kilohectares (kha)
**Reef_100**          Amount of other reef within 100km (kha)
-------------     ------------------------------------------------------------------------------


# Set model formula
Response variable is SpecRich
```{r}
div_isol_form <- bf(SpecRich ~ scale(Reef_1_Area) * scale(Reef_100) + (1|Site),  family = gaussian(link = "log"))
```

# What priors do we need?
```{r}
get_prior(div_isol_form, data = species_rich_asymp)
```

# What is the range of our data?
```{r}
species_rich_asymp |> 
  mutate(log_SR = log(SpecRich)) |> 
  group_by(Depth) |> 
  summarise(median = median(SpecRich), MAD = mad(SpecRich), 
            log_median = median(log_SR), log_MAD = mad(log_SR))
```


# Weakly informative priors
```{r}
div_isol_priors <- 
  prior(normal(3.2, 0.75), class = 'Intercept') +      
  prior(normal(0, 0.25), class = 'b') +
  prior(student_t(3, 0, 2.5) , class = 'sigma') 
 # prior(cauchy(0, 2), class = 'shape')
```


# Fit model
```{r}
div_isol_prioronlymod<- brm(formula = div_isol_form, 
            data = species_rich_asymp,
            prior = div_isol_priors,
            sample_prior = 'only',             
            chains = 3,
            cores = 3,             
            iter = 5000,
            warmup = 1000,  
            seed = 1234,
            thin = 4,
            silent = 0,            
            refresh = 0, 
            backend = "cmdstanr")
```



## Check
```{r}
div_isol_prioronlymod |> conditional_effects() |> plot(points = TRUE, ask = FALSE)
```



## Check
```{r}
div_isol_prioronlymod |> conditional_effects() |> plot(points = TRUE, ask = FALSE)
```




# Priors look good - include data
```{r}
div_isol_mod <- div_isol_prioronlymod |> update(sample_prior = 'yes', control = list(adapt_delta = 0.99, max_treedepth = 20),
                                           cores = 3, seed = 1234, refresh = 0)
```


# CHeck prior - posterior influence
```{r}
div_isol_mod |> SUYR_prior_and_posterior()
```


Looking good


# Chain diagnostics

## Trace plots
```{r traceplot}
div_isol_mod$fit |>  stan_trace()
```
## Autocorrelation
```{r}
div_isol_mod$fit |> stan_ac() 
```
## rhat
```{r}
div_isol_mod$fit |> stan_rhat() 
```
## Effectiveness of sampling
```{r}
div_isol_mod$fit |> stan_ess()
```


# Residuals
```{r}
resids <- make_brms_dharma_res(div_isol_mod, integerResponse = FALSE)
testUniformity(resids)
plotResiduals(resids, form = factor(rep(1, nrow(species_rich_asymp))))
plotResiduals(resids, quantreg = FALSE)
testDispersion(resids) # Dispersion plot (dispersion = variance/mean) - we assume mean and variance are the same, so we need to check
```


# Residuals look ok

## PP checks
```{r}
pp_check(div_isol_mod, ndraws = 250)
```

# Model looks good.
However, we also have the option of using a Gamma distribution

# Set model formula
```{r}
div_isol_form2 <- bf(SpecRich ~ scale(Reef_1_Area) * scale(Reef_100) + (1|Site),  family = Gamma(link = "log"))
```

# What priors
```{r}
get_prior(div_isol_form2, data = species_rich_asymp)
```

# Similar priors but now with one for shape
```{r}
div_isol_priors2 <- 
  prior(normal(3.75, 0.75), class = 'Intercept') +     
  prior(normal(0, 0.25), class = 'b') +
  prior(cauchy(0, 2), class = 'shape')

```


# Fit model
```{r}
div_isol_prioronlymod2<- brm(formula = div_isol_form2, 
            data = species_rich_asymp,
            prior = div_isol_priors2,
            sample_prior = 'only',             
            chains = 3,
            cores = 3,             
            iter = 5000,
            warmup = 1000,         
            thin = 4,
            silent = 0,            
            refresh = 0, 
            backend = "cmdstanr")
```


## Check
```{r}
div_isol_prioronlymod2 |> conditional_effects() |> plot(points = TRUE, ask = FALSE)
```


# Priors look good - include data
```{r}
div_isol_mod2 <- div_isol_prioronlymod2 |> update(sample_prior = 'yes', control = list(adapt_delta = 0.99, max_treedepth = 20),
                                           cores = 3, seed = 1234, refresh = 0)
```



# CHeck prior - posterior influence
```{r}
div_isol_mod2 |> SUYR_prior_and_posterior()
```


# Chain diagnostics

## Trace plots
```{r traceplot}
div_isol_mod2$fit |>  stan_trace()
```
## Autocorrelation
```{r}
div_isol_mod2$fit |> stan_ac() 
```
## rhat
```{r}
div_isol_mod2$fit |> stan_rhat() 
```
## Effectiveness of sampling
```{r}
div_isol_mod2$fit |> stan_ess()
```


# Residuals
```{r}
resids <- make_brms_dharma_res(div_isol_mod2, integerResponse = FALSE)
testUniformity(resids)
plotResiduals(resids, form = factor(rep(1, nrow(species_rich_asymp))))
plotResiduals(resids, quantreg = FALSE)
testDispersion(resids) # Dispersion plot (dispersion = variance/mean) - we assume mean and variance are the same, so we need to check
```



# Residuals look ok

## PP checks
```{r}
# Exponential
pp_check(div_isol_mod2, ndraws = 250)

# Logged
pred <- posterior_predict(div_isol_mod2)
bayesplot::ppc_dens_overlay(y = log1p(species_rich_asymp$SpecRich), 
                            yrep = log1p(pred[1:250,]))
```



# Compare models

```{r}
loo_compare(brms::loo(div_isol_mod),  # The log normal model
            brms::loo(div_isol_mod2)) # The gamma model
```

We'll take model 2 forward for further investigation

# Check R2
```{r}
bayes_R2(div_isol_mod)
bayes_R2(div_isol_mod2)
```



## Summary table
```{r}
div_isol_mod2 |> 
  as_draws_df() |>    # Extract draws
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  rhat, ess_bulk, ess_tail,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1))|> 
  as_tibble() |> 
  mutate(median = round(median, 3),
         lower = round(lower, 3),
         upper = round(upper, 3))
```






# Final model
```{r}
div_isol_mod_final <- div_isol_mod2
save(div_isol_mod_final, file = "data/Rdata/div_isol_model_final.Rdata")
```
