---
title: "Diversity modelling - Isolation"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
    css: kable.css
editor_options: 
  chunk_output_type: inline  
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Packages and other functions
```{r}
source('code/packages.R')
source('code/extras.R')
source('code/helper_functions.R')
```


# Load model and data
```{r}
load(file = "data/Rdata/species_rich_asymp.Rdata")
load("data/Rdata/div_biogeo_model_final.Rdata")
```

# Model investigation
```{r}
div_biogeo_mod_final |> conditional_effects() |> plot(points = TRUE, ask = FALSE) 
```


## Summary table
```{r}
div_biogeo_mod_final |> 
  gather_draws(`^b_.*`, regex=TRUE) |> 
  #filter(.variable != 'b_Intercept') |> 
  mutate(Value = exp(.value), .keep = "unused")  |> 
  group_by(.variable) |> 
  summarise(median_hdci(Value),
            Pl = mean(Value < 1),
            Pg = mean(Value > 1))
```

## Alternative
```{r}
div_biogeo_mod_final |> 
  as_draws_df() |>    # Extract draws
  #select(1:4)  |> 
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  rhat, ess_bulk, ess_tail,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1)) |> 
  as_tibble()  |> 
  mutate(across(where(is.numeric), round, 4))
```

# Summary

The only variable with some evidence of an effect is dGBR (P=0.98)

Let's do a pairwise comparison at max and min dGBR

```{r}
predsum |> 
  summarise(min_dGBR = min(dGBR),
            max_dGBR = max(dGBR))
```

# New data for emmeans
```{r}
dgbr_newdata <- data.frame(dGBR = c(48, 355))
```

# Basic emmeans
```{r}
div_biogeo_mod_final %>% 
  emmeans(~ dGBR, at = dgbr_newdata, type = "response") 
```

So, on average 36 predator species per site at reefs closest to the GBR and 14 at those furthest away. Should have a diff of about 22 in the pairwise comparison

```{r}
div_biogeo_mod_final |> 
  emmeans(~ dGBR, at = dgbr_newdata, type = "link")  |> 
  regrid() |> 
  pairs(reverse = FALSE)
```

Yep, but looks "non-significant" in frequentist terms

# As a ratio:
```{r}
div_biogeo_mod_final |> 
  emmeans(~ dGBR, at = dgbr_newdata, type = "response")  |> 
  pairs(reverse = FALSE)
```
2.6 times more predators at near-GBR reefs



With probability of a difference

```{r}
div_biogeo_mod_final |> 
  emmeans(~ dGBR, at = dgbr_newdata, type = "link")  |> 
  pairs(reverse = FALSE) |> 
  gather_emmeans_draws() |> 
  summarise(median_hdci(exp(.value)), Pg = mean(.value >0))
```

98% probability of an increase in predators going from most to least offshore reefs.

Diff in terms of numbers of species
```{r}
div_biogeo_mod_final |> 
  emmeans(~ dGBR, at = dgbr_newdata, type = "link")  |> 
  regrid() |> 
  pairs(reverse = FALSE) |> 
  gather_emmeans_draws() |> 
  summarise(median_hdci(.value), Pg = mean(.value >0))
```

22

# Figures


# dGBR trend

# Make new data grid
```{r}
dGBR_grid <- predsum |> 
  with(list(dGBR = seq_range(dGBR, n = 100)))
```

# Plot trend
```{r, fig.width=100/25.4, fig.height=50/26.5,}
dGBR_div_trend <- 
div_biogeo_mod_final |> 
  emmeans(~ dGBR, epred = TRUE,
           at = dGBR_grid) |> 
  gather_emmeans_draws() |> 
  ggplot(aes(x = dGBR, y = .value)) +
  stat_lineribbon(colour = "black", fill = "#FFCD00", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.5) +
  stat_lineribbon(colour = "black", fill = "#FFCD00", size = 0.5, point_interval = "median_hdci", .width = 0.8, alpha = 1) +
  scale_y_continuous("Sp. richness (.site)", expand = c(0.05,0)) +
  scale_x_continuous("Distance from GBR (km)") + 
  theme_bjc() +
  theme(axis.title.x = element_text(colour = "black", size = 10),
        axis.title.y = element_text(colour = "black", size = 10),
        axis.text = element_text(colour = "black", size = 9),
        legend.position = "none")

dGBR_div_trend
```


# Probability distribution of effect size
```{r fig.width = 50/25.4, fig.height=55/25.4}
dGBR_div_effect <- 
div_biogeo_mod_final |> 
  gather_draws(`^b_.*`, regex=TRUE) |>   # Need to use gather_draws here if you want to use the regex
  filter(.variable == 'b_scaledGBR') |>
  ungroup() |> 
  ggplot(aes(x = .value)) + 
  stat_halfeye(fill = "#FFCD00", .width = c(0.8, 0.95), point_interval = "median_hdi") +
  geom_vline(xintercept = 0, lty = 11) +
  annotate(geom = "text", label = "P=0.98", x = -0.6, y = 0.8, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  scale_y_discrete(expand = c(0.05,0)) +
  scale_x_continuous("Scaled effect distribution", breaks = c(-0.5, 0)) +
  theme_bjc() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(colour = "black", size = 10),
        axis.text.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 9),
        legend.position = "none")

dGBR_div_effect
```


# Combine
```{r fig.width = 150/25.4, fig.height=55/25.4}
dGBR_div_fig <- 
  cowplot::plot_grid(dGBR_div_trend, dGBR_div_effect, 
                     nrow = 1, labels = c("a", "b"), rel_widths = c(1.2, 0.6), label_size = 10, align = "h") +
  theme(plot.background = element_rect(color = "black"))
dGBR_div_fig
```


# Save
```{r}
ggsave(filename = "output/dGBR_div_fig.pdf", plot = dGBR_div_fig, height = 60, width = 152, device = cairo_pdf, units = "mm", dpi = 300)
```




# dCT - NO EFFECT

# Make new data grid
```{r}
dCT_grid <- predsum |> 
  with(list(dCT = seq_range(dCT, n = 100)))
```


# Trend ribbon plot
```{r, fig.width=100/25.4, fig.height=50/26.5,}
dCT_div_trend <- 
div_biogeo_mod_final |> 
  emmeans(~ dCT, type = 'link', 
          epred = TRUE, 
          at = dCT_grid) |> 
  regrid()  |> 
  gather_emmeans_draws()  |> 
  as_tibble() |> 
  ggplot(aes(x = dCT, y = .value)) +
  stat_lineribbon(colour = "black", fill = "#C65894", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.5) +
  stat_lineribbon(colour = "black", fill = "#C65894", size = 0.5, point_interval = "median_hdci", .width = 0.8, alpha = 1) +
  scale_x_continuous("Distance from Coral Triangle (km)") +
  scale_y_continuous("Sp. richness (.site)", expand = c(0.05,0)) +
  theme_bjc() +
  theme(axis.title.x = element_text(colour = "black", size = 10),
        axis.title.y = element_text(colour = "black", size = 10),
        axis.text = element_text(colour = "black", size = 9),
        legend.position = "none")

dCT_div_trend
```


# Effect size density plot
```{r fig.width=56/25.4, fig.height=60/25.4}
dCT_div_effect <- 
  div_biogeo_mod_final |> 
  gather_draws(`^b_.*`, regex=TRUE)  |> 
  filter(.variable == "b_scaledCT")   |> 
  ungroup() %>% 
  ggplot(aes(x = .value)) +
  stat_halfeye(.width = c(0.8, 0.95), fill = "#C65894", point_interval = "median_hdi") +
  annotate(geom = "text", label = "P=0.69", x = 0.3, y = 1, colour = "black", size = 3.5, family = "Arial", hjust = 0) +
  scale_x_continuous("Scaled effect distribution", breaks = c(-0.5, 0, 0.5)) +
  scale_y_discrete(expand = c(0.05,0)) +
  geom_vline(xintercept = 0, lty = 11) +
  theme_bjc() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(colour = "black", size = 10),
        axis.text.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 9),
        legend.position = "none")

dCT_div_effect
```

```{r fig.width = 150/25.4, fig.height=55/25.4}
dCT_div_fig <- 
  cowplot::plot_grid(dCT_div_trend, dCT_div_effect, 
                     nrow = 1, labels = c("a", "b"), rel_widths = c(1.2, 0.6), label_size = 10, align = "h") +
  theme(plot.background = element_rect(color = "black"))

dCT_div_fig
```


# Save
```{r}
ggsave(filename = "output/dCT_div_fig.pdf", plot = dCT_div_fig, height = 60, width = 152, device = cairo_pdf, units = "mm", dpi = 300)
```


```{r}
# END #
```


