---
title: "Diversity modelling - Isolation"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
    css: kable.css
editor_options: 
  chunk_output_type: inline  
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```

# Packages and extras
```{r}
source('code/packages.R')
source('code/helper_functions.R')
source('code/extras.R')
```

# Load model and data
```{r}
load(file = "data/Rdata/species_rich_asymp.Rdata")
load("data/Rdata/div_isol_model_final.Rdata")
load("data/Rdata/predsum.Rdata")
```

# Model investigation
```{r}
div_isol_mod_final |> conditional_effects() |> plot(points = FALSE, ask = FALSE)
```


```{r}
div_isol_mod_final |> conditional_effects() |> plot(points = TRUE, ask = FALSE)
```

Eyeballing this: no interaction, no effect of reef within 100km. Negative effect of reef area - more detail:

## Summary table
```{r}
div_isol_mod_final |> 
  gather_draws(`^b_.*`, regex=TRUE) |> 
  #filter(.variable != 'b_Intercept') |> 
  mutate(Value = exp(.value), .keep = "unused")  |> 
  group_by(.variable) |> 
  summarise(median_hdci(Value),
            Pl = mean(Value < 1),
            Pg = mean(Value > 1))
```

## Alternative
```{r}
div_isol_mod_final |> 
  as_draws_df() |>    # Extract draws
  #select(1:4)  |> 
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  rhat, ess_bulk, ess_tail,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1)) |> 
  as_tibble()  |> 
  mutate(across(where(is.numeric), round, 4))
```

- At the average reef size and isolation (the intercept) the median number of predatory fish species is 17 per site 95% HDI [12-22]
- Evidence of a 20% decline in species richness for every one unit increase in reef size (Pl = 0.94)
- No evidence of an effect of reef within 100km
- There is no evidence of an interaction



# emmeans and contrasts
Will compare abundance at max and min values of both predictors

```{r}
predsum |> 
  summarise(min_Reef_1_Area = min(Reef_1_Area),
            max_Reef_1_Area = max(Reef_1_Area),
            min_Reef100 = min(Reef_100),
            max_Reef100 = max(Reef_100))
```

# Make dataframe with data points:
```{r}
isol_newdata <- data.frame(Reef_1_Area = c(1.25, 26.4),
                      Reef_100 = c(1, 32.6))
```


# Reef Area

## Check amounts at each end to help guide interpretation
```{r}
div_isol_mod_final |> 
  emmeans(~ Reef_1_Area, at = isol_newdata, type = "response") 
```

So almost twice the species richness at the smallest reefs compared to the largest. Check with pair-wise comparisons

```{r}
div_isol_mod_final |> 
  emmeans(~ Reef_1_Area, at = isol_newdata, type = "link")  |> 
  regrid() |> 
  pairs(reverse = FALSE)
```
 contrast                          estimate lower.HPD upper.HPD
 Reef_1_Area1.25 - Reef_1_Area26.4     9.71     -2.08      24.9

Results are averaged over the levels of: Reef_100 
Point estimate displayed: median 
HPD interval probability: 0.95 

## Nearly 10 more species per site, although there is some clearly some uncertainty around this

As a ratio:

```{r}
div_isol_mod_final |> 
  emmeans(~ Reef_1_Area, at = isol_newdata, type = "response") |> 
  pairs(reverse = FALSE)
```
67% more species on smallest reefs

# Exceedence probability
```{r}
div_isol_mod_final |> 
  emmeans(~ Reef_1_Area, at = isol_newdata, type = "link") |> 
  pairs(reverse = FALSE) |> 
  gather_emmeans_draws() |> 
  summarise(median_hdci(exp(.value)), Pg = mean(.value >0))
```


Evidence to support a 60% increase in predator species between the largest and smallest reef 


```{r}
div_isol_mod_final |> 
  emtrends(~ Reef_100, var = "Reef_1_Area") %>% 
  regrid() |> 
  as_tibble()
```



# Plots to illustrate effects

## Reef area - halfeye plot for effect distribution 

```{r fig.width = 70/25.4, fig.height=60/25.4}
div_reefarea_effect <- 
div_isol_mod_final |> 
  gather_draws(b_scaleReef_1_Area) |> # Option 1 for filtering out the draws you want
  #gather_draws(`^b_.*`, regex=TRUE) |>                 # Option 2 
  #filter(.variable != 'b_Intercept') |>
  #filter(.variable != 'b_zi_Intercept') |> 
 # filter(.variable != 'b_scaleReef_1_Area:scaleReef_100') |> 
  ggplot() + 
  stat_halfeye(aes(x=.value), fill = "#cc0000", .width = c(0.8, 0.95), point_interval = "median_hdi") +
  #ggridges::geom_density_ridges_gradient(aes(x=exp(.value),  y=.variable, fill = stat(x)), alpha = 0.4, colour = "black", quantile_lines = TRUE, quantiles = c(0.025, 0.5, 0.975)) +
  geom_vline(xintercept = 0, lty = 11) +
  annotate(geom = "text", label = "P=0.97", x = -0.5, y = 0.8, colour = "black", size = 3.5, family = "Arial") +
  scale_y_discrete(expand = c(0.05,0)) +
  scale_x_continuous("Scaled effect distribution", breaks = c(-0.5, 0, 0.5)) +
  theme_bjc() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(colour = "black", size = 10),
        axis.text.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 9),
        legend.position = "none")

div_reefarea_effect
```


# Trend

# emmeans for plot data
```{r}
reefarea_newdata <- with(species_rich_asymp, list(Reef_1_Area = seq(from = min(Reef_1_Area), to = max(Reef_1_Area), len = 100)))
```


# Reef area trend - using draws (to get 85 and 95% CI)
```{r fig.height=60/26.5, fig.width=110/25.4}
div_reefarea_trend <- 
div_isol_mod_final |> 
  emmeans(~ Reef_1_Area, epred = TRUE,
           at = reefarea_newdata) |> 
  gather_emmeans_draws() |> 
  ggplot(aes(x = Reef_1_Area, y = .value)) +
 # geom_point(data = predsum %>% filter(No_Preds != 0), aes(x = Reef_1_Area, y = No_Preds), fill = "#cc0000", colour = "#cc0000", pch = 21, alpha = 0.3, position = position_jitter()) +
  stat_lineribbon(colour = "black", fill = "#cc0000", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.5) +
  stat_lineribbon(colour = "black", fill = "#cc0000", size = 0.5, point_interval = "median_hdci", .width = 0.80, alpha = 1) +
  scale_x_continuous("Reef area (kha)") +
  scale_y_continuous("Species richness (.site)", expand = c(0.05,0)) +
  theme_bjc() +
  theme(axis.title.x = element_text(colour = "black", size = 10),
        axis.title.y = element_text(colour = "black", size = 10, hjust = -1), # This looks horrible on the single plot but reqd for the combined fig. see below
        axis.text = element_text(colour = "black", size = 9),
        legend.position = "none")
div_reefarea_trend
```

# Combine
```{r fig.width = 150/25.4, fig.height=55/25.4}
div_reefarea_plot <- 
  cowplot::plot_grid(div_reefarea_trend, div_reefarea_effect, 
                     nrow = 1, labels = c("a", "b"), rel_widths = c(1.2, 0.6), label_size = 10, align = "h") +
  theme(plot.background = element_rect(color = "black"))
div_reefarea_plot
```


# Save
```{r}
ggsave(filename = "output/reefarea_div_fig.pdf", plot = div_reefarea_plot, width = 150, height = 55, device = cairo_pdf, units = "mm", dpi = 300)
```


# Contrast: largest - smallest reef
```{r fig.width = 40/25.4, fig.height=50/25.4}
div_area_contrast <- div_isol_mod_final |> 
  emmeans(~ Reef_1_Area, at = newdata, type = "link") |> 
  pairs(reverse = FALSE) |> 
  gather_emmeans_draws() |> 
  summarise(median_hdci(exp(.value), .width = c(0.8, 0.95)),
            Pl = mean(.value < 0),
            Pg = mean(.value > 0)) 



reefarea_div_full_contrast <- 
ggplot() +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax, x = contrast), data = div_area_contrast |> filter(.width == 0.8), linewidth = 1.7) +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax, x = contrast), data = div_area_contrast |> filter(.width == 0.95), pch = 21, size = 1.2, fill = "#cc0000", linewidth = 0.7) +
  geom_hline(yintercept = 1, lty = 11) +
  theme_bjc() +
  scale_y_continuous("Contrast ratio (max-min)", limits = c(0, 6), breaks = c(0,2,4)) +
  scale_x_discrete("Diversity contrast") +
   theme(axis.title = element_text(colour = "black", size = 10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(colour = "black", size = 9),
        legend.position = "none")

reefarea_div_full_contrast
```

# Save
```{r}
ggsave(filename = "output/reefarea_div_full_contrast.pdf", plot = reefarea_div_full_contrast, width = 50, height = 55, device = cairo_pdf, units = "mm", dpi = 300)
```




# Reef within 100km



## Check amounts at each end to help guide interpretation
```{r}
div_isol_mod_final |> 
  emmeans(~ Reef_100, at = isol_newdata, type = "response") 
```


# Contrasts
```{r}
div_isol_mod_final |> 
  emmeans(~ Reef_100, at = isol_newdata, type = "response")  |> 
  regrid() |> 
  pairs(reverse = FALSE)
```


# Halfeye plot
```{r fig.width = 70/25.4, fig.height=60/25.4}
div_reef100_effect <- 
div_isol_mod_final |> 
  gather_draws(b_scaleReef_100) |> # Option 1 for filtering out the draws you want
  ungroup() |> 
  ggplot(aes(x=.value)) + 
  stat_halfeye(fill = "#009933", .width = c(0.80, 0.95), point_interval = "median_hdi") +
  geom_vline(xintercept = 0, lty = 11) +
  annotate(geom = "text", label = "P=0.91", x = -0.5, y = 0.9, colour = "black", size = 3.5, family = "Arial") +
  scale_y_discrete(expand = c(0.05,0)) +
  scale_x_continuous("Scaled effect distribution", breaks = c(-0.5, 0, 0.5)) +
  theme_bjc() +
  theme(axis.title.x = element_text(colour = "black", size = 10),
        axis.title.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 9),
        axis.text.y = element_blank(),
        legend.position = "none")

div_reef100_effect
```


## Plot the trends

# emmeans for plot data
```{r}
reef100_newdata <- with(species_rich_asymp, list(Reef_100 = seq(from = min(Reef_100), to = max(Reef_100), len = 100)))
```


# Reef 100km trend - using draws
```{r fig.height=60/26.5, fig.width=110/25.4}
div_reef100_trend <- 
div_isol_mod_final |> 
  emmeans(~ Reef_100, epred = TRUE,
           at = reef100_newdata) |> 
  gather_emmeans_draws() |> 
  ggplot(aes(x = Reef_100, y = .value)) +
  #geom_point(data = predsum %>% filter(No_Preds != 0), aes(x = Reef_100, y = No_Preds), fill = "#66cc99", colour = "#336633", pch = 21, alpha = 0.3, position = position_jitter()) +
  stat_lineribbon(colour = "black", fill = "#009933",  size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.5) +
  stat_lineribbon(colour = "black", fill = "#009933", size = 0.5, point_interval = "median_hdci", .width = 0.85, alpha = 1) +
  scale_x_continuous("Reef within 100km (kha)") +
  scale_y_continuous("Species richness (.site)", expand = c(0.05,0)) +
  theme_bjc() +
  theme(axis.title.x = element_text(colour = "black", size = 10),
        axis.title.y = element_text(colour = "black", size = 10, hjust = -0.5),
        axis.text = element_text(colour = "black", size = 9),
        legend.position = "none")

div_reef100_trend
```



```{r fig.width = 150/25.4, fig.height=55/25.4}
div_reef100_plot <- 
  cowplot::plot_grid(div_reef100_trend, div_reef100_effect, 
                     nrow = 1, labels = c("a", "b"), rel_widths = c(1.2, 0.6), label_size = 10, align = "h") +
    theme(plot.background = element_rect(color = "black"))

div_reef100_plot
```



# Save
```{r}
ggsave(filename = "output/reef100_div_fig.pdf", plot = div_reef100_plot, width = 150, height = 55, device = cairo_pdf, units = "mm", dpi = 300)
```




# Contrast: most to least spatially isolated reef reef
```{r fig.width = 40/25.4, fig.height=50/25.4}
reef100_div_contrast <- div_isol_mod_final |> 
  emmeans(~ Reef_100, at = isol_newdata, type = "link") |> 
  pairs(reverse = FALSE) |> 
  gather_emmeans_draws() |> 
  summarise(median_hdci(exp(.value), .width = c(0.8, 0.95)), 
            Pl = mean(.value < 0),
            Pg = mean(.value > 0))
  
abun_isol_mod_final |> 
  emmeans(~ Reef_100, at = newdata, type = "link") |> 
  pairs(reverse = FALSE) |> 
  regrid() |> 
  gather_emmeans_draws() |> 
  summarise(median_hdci(exp(.value)), Pg = mean(.value >0))
```


```{r fig.width = 40/25.4, fig.height=50/25.4}
reef100_div_full_contrast <- 
ggplot() +
  geom_hline(yintercept = 1, lty = 11) +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax, x = contrast), data = reef100_div_contrast |> filter(.width == 0.8), linewidth = 1.7) +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax, x = contrast), data = reef100_div_contrast |> filter(.width == 0.95), pch = 22, size = 1.2, fill = "#009933", linewidth = 0.7) +
  theme_bjc() +
  scale_y_continuous("Contrast ratio (max-min)") +
  scale_x_discrete("Reef isolation contrast") +
   theme(axis.title = element_text(colour = "black", size = 10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(colour = "black", size = 9),
        legend.position = "none")

reef100_div_full_contrast
```

# Save
```{r}
ggsave(filename = "output/reef100_div_full_contrast.pdf", plot = reef100_div_full_contrast, width = 50, height = 55, device = cairo_pdf, units = "mm", dpi = 300)
```



```{r}
# END #
```










