---
title: "Diversity modelling - Isolation"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
    css: kable.css
editor_options: 
  chunk_output_type: inline  
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


# Packages and other functions
```{r}
source('code/packages.R')
source('code/extras.R')
source('code/helper_functions.R')
```


# Load model and data
```{r}
load(file = "data/Rdata/species_rich_asymp.Rdata")
load("data/Rdata/div_biogeo_model_final.Rdata")
```

# Model investigation
```{r}
div_biogeo_mod_final |> conditional_effects() |> plot(points = TRUE, ask = FALSE)
```


## Summary table
```{r}
div_biogeo_mod_final |> 
  gather_draws(`^b_.*`, regex=TRUE) |> 
  #filter(.variable != 'b_Intercept') |> 
  mutate(Value = exp(.value), .keep = "unused")  |> 
  group_by(.variable) |> 
  summarise(median_hdci(Value),
            Pl = mean(Value < 1),
            Pg = mean(Value > 1))
```

## Alternative
```{r}
div_biogeo_mod_final |> 
  as_draws_df() |>    # Extract draws
  #select(1:4)  |> 
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1)) |> 
  as_tibble()  |> 
  mutate(across(where(is.numeric), round, 4))
```

# Summary

- At the average depth, dCT and dGBR (the intercept) the median number of predatory fish species is 14 per transect 95% HDI [9-19]

# Interaction - dGBR * Depth
No effect of distance from GBR on species richness at any depth

# Interaction - dCT * Depth
In the shallows, species richness increased with distance from the Coral Triangle, although this was highly variable





# Effect distributions

# Distribution of dCT effect
```{r fig.width = 70/25.4, fig.height=60/25.4}
div_biogeo_mod_final |> 
  gather_draws(b_scaledCT) |> # Option 1 for filtering out the draws you want
  #gather_draws(`^b_.*`, regex=TRUE) |>                 # Option 2 
  #filter(.variable != 'b_Intercept') |>
  #filter(.variable != 'b_zi_Intercept') |> 
 # filter(.variable != 'b_scaleReef_1_Area:scaleReef_100') |> 
  ggplot() + 
  stat_halfeye(aes(x=exp(.value),  y=.variable, fill = .variable), .width = c(0.8, 0.95), point_interval = "median_hdi", alpha = 0.8) +
  #ggridges::geom_density_ridges_gradient(aes(x=exp(.value),  y=.variable, fill = stat(x)), alpha = 0.4, colour = "black", quantile_lines = TRUE, quantiles = c(0.025, 0.5, 0.975)) +
  geom_vline(xintercept = 1, linetype = 'dashed') +
  scale_fill_manual("Effect", values = c("#fa2600")) +
  scale_y_discrete(position = "right", labels = c("Reef\nsize", "Reef\nisol"), expand = c(0.01, 0.01)) +
  scale_x_continuous("Effect distribution", trans = scales::log2_trans(), expand = c(0.15,0.1)) +
  theme_minimal() +
  theme(axis.title.y.right = element_blank(),
        axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        strip.text = element_text(colour = "black", size = 9),
        plot.title = element_text(size = 10, hjust = 0.5),
        axis.text.y.right = element_blank(),
        legend.position = "none")
```



# Distribution of dGBR effect
```{r fig.width = 70/25.4, fig.height=60/25.4}
div_biogeo_mod_final |> 
  gather_draws(b_scaledGBR) |> # Option 1 for filtering out the draws you want
  ggplot() + 
  stat_halfeye(aes(x=exp(.value),  y=.variable, fill = .variable), .width = c(0.8, 0.95), point_interval = "median_hdi", alpha = 0.8) +
  #ggridges::geom_density_ridges_gradient(aes(x=exp(.value),  y=.variable, fill = stat(x)), alpha = 0.4, colour = "black", quantile_lines = TRUE, quantiles = c(0.025, 0.5, 0.975)) +
  geom_vline(xintercept = 1, linetype = 'dashed') +
  scale_fill_manual("Effect", values = c("#006ad1")) +
  scale_y_discrete(position = "right", labels = c("Reef 100km"), expand = c(0.05,0)) +
  scale_x_continuous("Effect distribution", trans = scales::log2_trans()) +
  theme_minimal() +
  theme(axis.title.y.right = element_blank(),
        axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        strip.text = element_text(colour = "black", size = 9),
        plot.title = element_text(size = 10, hjust = 0.5),
        axis.text.y.right = element_blank(),
        legend.position = "none")
```



## Extract range
```{r}
dCT_range <- species_rich_asymp %>% 
  with(list(dCT = seq_range(dCT, n = 100)))
```


## Plot
```{r fig.height=170/26.5, fig.width=70/25.4}
div_dCTdepth_trend <- 
  div_biogeo_mod_final |> 
  emmeans(~ dCT | Depth, var = "dCT",
           #regrid = "response",  type = "response",
          epred = TRUE,
           at = dCT_range) %>% 
  gather_emmeans_draws() %>% 
  as_tibble() %>% 
  ggplot(aes(x = dCT, y = .value)) +
  #stat_lineribbon(colour = "black", fill = "#ccccff", size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.6) +
  #stat_lineribbon(colour = "black", fill = "#000099", size = 0.5, point_interval = "median_hdci", .width = 0.85, alpha = 0.9) +
  stat_lineribbon(aes(fill = Depth), colour = "black",  size = 0.5, point_interval = "median_hdci", .width = 0.95, alpha = 0.6) +
  stat_lineribbon(aes(fill = Depth), colour = "black", size = 0.5, point_interval = "median_hdci", .width = 0.85, alpha = 0.9) +
  facet_wrap(~Depth, ncol = 1, scales = "free") +
  scale_fill_manual(values = websafe_bluepal) +
  scale_y_continuous(expand = c(0.1, 0)) +
  labs(x = "Dist. Coral Triangle (km)", 
       y = "Speceies richness (.site)",
       title = "") +
  theme_minimal() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        #strip.text = element_text(colour = "black", size = 9),
        strip.text = element_blank(),
        legend.position = 'none')

div_dCTdepth_trend
```


# Save
```{r}
ggsave(filename = "output/div_dCTdepth_fig.pdf", plot = div_dCTdepth_trend, height = 160, width = 70, device = cairo_pdf, units = "mm", dpi = 300)
```




## Extra stuff/junk ##

# Plot

```{r}
species_rich_asymp  %>% 
  select(Reef_1, dCT) %>% 
  distinct() %>% 
  summarise(median = median(dCT), Q1 = quantile(dCT, probs = 0.25), Q4 = quantile(dCT, probs = 0.95),
            CI_05 = quantile(dCT, probs = 0.05), CI_95 = quantile(dCT, probs = 0.99),
            mean = mean(dCT))
```


# Make new data grid
```{r}
biogeogrid <- species_rich_asymp |> 
  with(list(Depth = levels(Depth),
            dCT = c(50, 600, 1000)))
```


# dCT * Depth
```{r fig.height=120/26.5, fig.width=100/25.4}
div_biogeo_mod_final |> 
  emmeans(~ Depth|dCT, at = biogeogrid, type = 'response') |> 
  as_tibble() |> 
  ggplot(aes(x = Depth, y = response)) +
  geom_line(aes(group = 1)) +
  geom_pointrange(aes(ymin = lower.HPD, ymax = upper.HPD, fill = Depth), colour = "black", pch = 21, size = 1) +
  scale_fill_manual(values = c("#006ad1", "#1e90ff", "#6bb6ff")) +
  #scale_colour_manual(values = c("#006ad1", "#1e90ff", "#6bb6ff")) +
  theme_minimal() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        strip.text = element_text(colour = "black", size = 9),
        plot.title = element_text(size = 10, hjust = 0.5),
        legend.position = "none") +
  facet_wrap(~dCT, nrow = 3, strip.position = "right")
```




## Plot the trends

# emmeans for plot data

```{r}
newdata <- with(species_rich_asymp,
                list(Reef_1_Area = seq(from = min(Reef_1_Area), to = max(Reef_1_Area), len = 20),
                     Reef_100 = seq(from = min(Reef_100), to = max(Reef_100), len = 20)))

```

# Reef area trend
```{r fig.width = 100/25.4, fig.height=60/25.4}
reefarea_trend <- 
  div_biogeo_mod_final |> 
  emmeans(~Reef_1_Area, type = "response",
          at = newdata)  |> 
  as_tibble() |> 
  ggplot() +
  geom_ribbon(aes(x = Reef_1_Area, ymin = lower.HPD, ymax = upper.HPD), fill = "#fa2600", alpha = 0.8) +
  geom_line(aes(x = Reef_1_Area, y = response)) +
  labs(x = "Reef size (kha)", y = "Species richness (.250m2)") +
  scale_y_continuous() +
  theme_minimal() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        strip.text = element_text(colour = "black", size = 9),
        plot.title = element_text(size = 10, hjust = 0.5),
        legend.position = "none") 
reefarea_trend
```


# Reef 100km trend
```{r fig.width = 100/25.4, fig.height=60/25.4}
reef100_trend <- 
div_biogeo_mod_final |> 
  emmeans(~Reef_100, type = "response",
          at = newdata)  |> 
  as_tibble() |> 
  ggplot() +
  geom_ribbon(aes(x = Reef_100, ymin = lower.HPD, ymax = upper.HPD), fill = "#006ad1", alpha = 0.8) +
  geom_line(aes(x = Reef_100, y = response)) +
  labs(x = "Reef w/in 100km (kha)", y = "Species richness (.250m2)") +
  scale_y_continuous() +
  theme_minimal() +
  theme(axis.text = element_text(colour = "black", size = 9),
        axis.title = element_text(colour = "black", size = 10),
        strip.text = element_text(colour = "black", size = 9),
        plot.title = element_text(size = 10, hjust = 0.5),
        legend.position = "none") 
reef100_trend
```


# Combine
```{r fig.width = 170/25.4, fig.height=60/25.4}
reefarea_plot <- 
  cowplot::plot_grid(reefarea_trend, reefarea_effect, rel_widths = c(1, 0.7))
reefarea_plot
```



```{r fig.width = 170/25.4, fig.height=60/25.4}
reef100_plot <- 
  cowplot::plot_grid(reef100_trend, reef100_effect, rel_widths = c(1, 0.7))
reef100_plot
```


















