---
title: "Diversity modelling - biogeographic drivers"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
# Set options, housekeeping
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```

```{r}
source('code/packages.R')
source('code/helper_functions.R')
```


# We'll start with the aysymptotic species richness estimates from iNEXT.
```{r}
load(file = "data/Rdata/species_rich_asymp.Rdata")
```

```{r}
glimpse(species_rich_asymp)
```


As with the abundance modelling, we are going to examine the effect of the dCT and dGBR:


# Set model formula
Response variable is SpecRich
```{r}
div_biogeo_form <- bf(SpecRich ~ scale(dCT) * scale(dGBR) + (1|Reef_1),  family = gaussian(link = "log"))
```

# What priors do we need?
```{r}
get_prior(div_biogeo_form, data = species_rich_asymp)
```

# What is the range of our data?
```{r}
species_rich_asymp |> 
  mutate(log_SR = log(SpecRich)) |> 
 # group_by(Depth) |> 
  summarise(median = median(SpecRich), MAD = mad(SpecRich), 
            log_median = median(log_SR), log_MAD = mad(log_SR))
```


# Weakly informative priors
Need to go a little wider than for the isolation models
```{r}
div_biogeo_priors <- 
  prior(normal(2.7, 0.5), class = 'Intercept') +      
  prior(normal(0, 0.5), class = 'b') +
  prior(student_t(3, 0, 2.5) , class = 'sigma') 
 # prior(cauchy(0, 2), class = 'shape')

div_biogeo_priors <- 
prior(normal(3.5, 0.75), class = 'Intercept') +     
  prior(normal(0, 0.2), class = 'b') +
  prior(student_t(3, 0, 0.75) , class = 'sigma') 



```


# Fit model
```{r}
div_biogeo_prioronlymod<- brm(formula = div_biogeo_form, 
            data = species_rich_asymp,
            prior = div_biogeo_priors,
            sample_prior = 'only',             
            chains = 3,
            cores = 3,             
            iter = 5000,
            warmup = 1000,         
            thin = 4,
            seed = 1234,
            silent = 0,            
            refresh = 0, 
            backend = "cmdstanr")
```



## Check
```{r}
div_biogeo_prioronlymod |> conditional_effects() |> plot(points = TRUE, ask = FALSE)
```



# Priors look good - include data
```{r}
div_biogeo_mod <- div_biogeo_prioronlymod |> update(sample_prior = 'yes', seed = 1234, control = list(adapt_delta = 0.99, max_treedepth = 20), refresh = 0)
```


# CHeck prior - posterior influence
```{r}
div_biogeo_mod |> SUYR_prior_and_posterior()
```


Looking good


# Chain diagnostics

## Trace plots
```{r traceplot}
div_biogeo_mod$fit |>  stan_trace()
```
## Autocorrelation
```{r}
div_biogeo_mod$fit |> stan_ac() 
```
## rhat
```{r}
div_biogeo_mod$fit |> stan_rhat() 
```
## Effectiveness of sampling
```{r}
div_biogeo_mod$fit |> stan_ess()
```


# Residuals
```{r}
resids <- make_brms_dharma_res(div_biogeo_mod, integerResponse = FALSE)
testUniformity(resids)
plotResiduals(resids, form = factor(rep(1, nrow(species_rich_asymp))))
plotResiduals(resids, quantreg = FALSE)
testDispersion(resids) # Dispersion plot (dispersion = variance/mean) - we assume mean and variance are the same, so we need to check
```


# Residuals look bad

## PP checks
```{r}
pp_check(div_biogeo_mod, ndraws = 250)
```

# And PP checks look weird also. Can use gamma:

# Set model formula
Response variable is SpecRich
```{r}
div_biogeo_form2 <- bf(SpecRich ~ scale(dCT) * scale(dGBR) + (1|Site),  family = Gamma(link = "log"))
```

# What priors do we need?
```{r}
get_prior(div_biogeo_form2, data = species_rich_asymp)
```

prior(normal(3, 0.5), class = 'Intercept') +      
  prior(normal(0, 0.4), class = 'b') +
  prior(student_t(3, 0, 2.5) , class = 'sigma') 

# Similar priors but now with one for shape
```{r}
div_biogeo_priors2 <- 
prior(normal(4, 0.8), class = 'Intercept') +     
  prior(normal(0, 0.25), class = 'b') +
  prior(cauchy(0, 2), class = 'shape')
```


# Fit model
```{r}
div_biogeo_prioronlymod2<- brm(formula = div_biogeo_form2, 
            data = species_rich_asymp,
            prior = div_biogeo_priors2,
            sample_prior = 'only',             
            chains = 3,
            cores = 3,             
            iter = 5000,
            warmup = 1000,         
            thin = 4,
            silent = 0,            
            refresh = 0, 
            backend = "cmdstanr")
```


## Check
```{r}
div_biogeo_prioronlymod2 |> conditional_effects() |> plot(points = TRUE, ask = FALSE)
```


# Priors look good - include data
```{r}
div_biogeo_mod2 <- div_biogeo_prioronlymod2 |> update(sample_prior = 'yes', seed = 1234, control = list(adapt_delta = 0.99, max_treedepth = 20), refresh = 0)
```



# CHeck prior - posterior influence
```{r}
div_biogeo_mod2 |> SUYR_prior_and_posterior()
```


# Chain diagnostics

## Trace plots
```{r traceplot}
div_biogeo_mod2$fit |>  stan_trace()
```
## Autocorrelation
```{r}
div_biogeo_mod2$fit |> stan_ac() 
```
## rhat
```{r}
div_biogeo_mod2$fit |> stan_rhat() 
```
## Effectiveness of sampling
```{r}
div_biogeo_mod2$fit |> stan_ess()
```


# Residuals
```{r}
resids <- make_brms_dharma_res(div_biogeo_mod2, integerResponse = FALSE)
testUniformity(resids)
plotResiduals(resids, form = factor(rep(1, nrow(species_rich_asymp))))
plotResiduals(resids, quantreg = FALSE)
testDispersion(resids) # Dispersion plot (dispersion = variance/mean) - we assume mean and variance are the same, so we need to check
```



# Residuals look ok

## PP checks
```{r}
# Exponential
pp_check(div_biogeo_mod2, ndraws = 250)

# Logged
pred <- posterior_predict(div_biogeo_mod2)
bayesplot::ppc_dens_overlay(y = log1p(species_rich_asymp$SpecRich), 
                            yrep = log1p(pred[1:200,]))
```



# Compare models

```{r}
loo_compare(brms::loo(div_biogeo_mod),  # The log normal model
            brms::loo(div_biogeo_mod2)) # The gamma model
```

We'll take model 2 forward for further investigation

# Summary

# Figure
```{r}
div_biogeo_mod2 |> conditional_effects() |> plot(points = FALSE, ask = FALSE)

```


Looks like:
No interaction
No effect of dCT
Neg effect of dGBR


## Table
```{r}
div_biogeo_mod2 |> 
  as_draws_df() |>    # Extract draws
  exp() |>            # Back transform
  summarise_draws(median,
                  HDInterval::hdi,
                  rhat, ess_bulk, ess_tail,
                  Pl = ~mean(.x <1),                       # Exceedence prob - on exp scale it's sig if doesn't cross 1
                  Pg = ~mean(.x >1))|> 
  as_tibble() |> 
  mutate(median = round(median, 3),
         lower = round(lower, 3),
         upper = round(upper, 3))
```

## Yep:
No interaction
No effect of dCT
Negative effect of dGBR P=0.98


# Check R2
```{r}
#bayes_R2(div_biogeo_mod)
bayes_R2(div_biogeo_mod2)
```


# Final model
```{r}
div_biogeo_mod_final <- div_biogeo_mod2
#save(div_biogeo_mod_final, file = "data/Rdata/div_biogeo_model_final.Rdata")
```
