---
title: "Predator diversity - initial wrangling and EDA"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	include = FALSE,
	message = FALSE,
	warning = FALSE)
```

# Housekeeping and packages
```{r include=FALSE}
#rm(list=ls()) # Clear out environment if necessary
getwd()
source("code/packages.R")
```

# Import predator data
```{r message=FALSE, warning=FALSE, include=FALSE}
source("code/wrangling/diw_rawdata.R")
```


# Notes
This EDA makes it evident that we don't really have enough information at transect level to make any inferences about differences in diversity at any spatial scale. We will probably need to pool the data at the next level up, which is Site * Depth.

```{r}
preds %>% 
  select(Site_Depth) %>% 
  distinct()
```
So we have 79 Site Depth combinations vs 538 transects. However, many of those transects are 0 observations (or 1s, but these are also problematic)


# Summarise spp counts by transect
```{r}
pred.count <- preds %>% 
  group_by(T_ID, Binomial) %>% # Group including family for use in taxa_nos analysis later
  dplyr::summarise(Count=sum(Number))
head(pred.count)
```

# Then pivot into wide diversity df
```{r}
pred.div <- pred.count %>%
  group_by(T_ID) %>% 
  pivot_wider(names_from = Binomial, values_from=Count, values_fill = 0)
```


# Add back in metadata and calculate species richness and Shannon by transect
```{r}
pred_diversity <-  rmetadata %>% 
  full_join(pred.div %>% 
      as.data.frame() |> 
      mutate(SpecNo = specnumber(pred.div[-1]), 
             Shannon = diversity(pred.div[-1])) %>% 
      select(T_ID, SpecNo, Shannon))
```


# So how many transects are 0s?
```{r}
pred_diversity %>% 
  filter(SpecNo == 0)
```


158 transects have 0 observations of predators! That's quite a lot 158/538 = nearly 30% of all transects.

At site-depth level it may be a different story

```{r}
pred_diversity %>% 
  group_by(Site_Depth) %>% 
  summarise(SpecNo = sum(SpecNo)) %>% 
  arrange(SpecNo)

```


6 zeros and 3 ones and 6 twos. That's a bit better.

What about sampling effort? We obviously haven't sampled every site - depth evenly? Rarefaction can probably handle some of that, as long as we leave the low digit entries out of the equation.

Let's check sampling effort

```{r}
pred_diversity %>% 
  group_by(Site_Depth) %>% 
  summarise(Number = n()) %>% 
  arrange(Number)
```







# EDA

# Reef scale differences?
```{r fig.width = 150/25.4, fig.height = 90/25.4}
pred_diversity %>% ggplot() +
  geom_boxplot(aes(x = Reef_1, y = SpecNo))
```


# Reef size
```{r}
pred_diversity %>% 
  ggplot(aes(x = Reef_1_Area, y = SpecNo)) +
  geom_point() +
  geom_smooth(method = 'glm') +
  theme_minimal()
```
Looks like there might be a decline in diversity with increasing reef size!?

# Spatial isolation
```{r}
pred_diversity %>% 
  ggplot(aes(x = Reef_100, y = SpecNo))+
  geom_point() +
  geom_smooth(method = 'glm') +
  theme_minimal()
```

Not much of a relationship...


# Aspect

```{r fig.width = 150/25.4, fig.height = 90/25.4}
pred_diversity %>% ggplot() +
  geom_boxplot(aes(x = Aspect_descriptive, y = SpecNo))
```


```{r fig.width = 150/25.4, fig.height = 90/25.4}
pred_diversity %>% ggplot() +
  geom_violin(aes(x = Situation, y = SpecNo))
```


```{r fig.width = 150/25.4, fig.height = 90/25.4}
pred_diversity %>% ggplot() +
  geom_violin(aes(x = Habitat, y = SpecNo))

```



# Habitat complexity
```{r fig.width = 150/25.4, fig.height = 90/25.4}
pred_diversity %>% ggplot()+
  geom_point(aes(x = Benthic_Complexity, y = SpecNo))
```

# Depth
```{r fig.width = 150/25.4, fig.height = 90/25.4}
pred_diversity %>% ggplot()+
    geom_point(aes(x = Depth_m, y = SpecNo)) +
    geom_smooth(aes(x = Depth_m, y = SpecNo)) +
  theme_minimal() +
  scale_y_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10))
```

# Depth - 3 zone:
```{r fig.width = 150/25.4, fig.height = 90/25.4}
pred_diversity %>% ggplot()+
    geom_boxplot(aes(x = Depth, y = SpecNo)) +
  scale_y_log10() +
  theme_minimal() 
```





# Temperature
```{r fig.width = 150/25.4, fig.height = 90/25.4}
pred_diversity %>% ggplot()+
  geom_point(aes(x = Temperature, y = Shannon))
```

# Latitude
```{r fig.width = 150/25.4, fig.height = 90/25.4}
pred_diversity %>% ggplot()+
  geom_point(aes(x = Site_lat, y = Shannon))
```

# Distance from CT
```{r}
pred_diversity %>% 
  ggplot(aes(x = dCT, y = Shannon)) +
  geom_point() +
  geom_smooth()
```






